```{r echo=FALSE}
# Description: 
# Set report parameters
Year = 2022 # Year of the Chisana comp count survey

```

---
title: "Results from the `r Year` Composition Count Survey of the Chisana Caribou
  Herd in Wrangell - St. Elias National Park and Preserve, Alaska"
author: ""
date: ""
output:
  html_document:
  toc: yes
---

```{r setup, include=FALSE}

# Suppress warnings
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Setup
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(knitr)
library(RODBC)
library(sf)
library(ggplot2)
library(tidyverse)
library(SightabilityModel)

# Restrain the number of decimal places
options(digits = 1) 

# WRST Shortcut
WRST = "Wrangell - St. Elias National Park and Preserve"

# Database connection string
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')

# Counters to increment the table and figure numbers as the report progresses.
TableCount = 1
FigureCount = 1

```

Kyle Cutting\
Wildlife Biologist\
National Park Service, `r WRST`\
8 Richardson Hwy Mile 106\
Copper Center, AK 99573

Scott D. Miller\
Information Technology Specialist/Data Manager\
National Park Service, Arctic Inventory and Monitoring Network\
240 W. 5th Ave.\
Anchorage, AK 99501

# Introduction

# Methods

Table X. Counts.
```{r echo=FALSE}

# Dates
Sql = paste("-- background
SELECT        [Start date], [End date], [Survey days], [Search areas]
FROM            Summary_Census
WHERE (Herd = 'Chisana') AND (Year = ",Year,")",sep="")

data = sqlQuery(Connection,Sql)
data_t = t(data)
colnames(data_t) = ''
kable(data_t)
```

## Distribution Map
```{r echo=FALSE}

# Get the spatial data for comp and population surveys
Sql = paste("SELECT        SurveyName, Year, Herd, SurveyType, SearchArea, GroupNumber, SightingDate, Lat, Lon
, IsNull([In],0) as [In]
, IsNull(Seen,0) as Seen
, IsNull(Marked,0) as Marked
FROM            CaribouGroups
WHERE (Year = ",Year,") And Herd='Chisana' And Lat > 0",sep="")
data = sqlQuery(Connection,Sql)

# Convert the data into a spatial data frame
spdata <- sf::st_as_sf(data, coords = c('Lon','Lat'))

# Set the coordinate system to GCS Lat\Lon WGS1984
st_crs(spdata) <- st_crs(4326)

# Park boundaries (geojson) are at https://irma.nps.gov/DataStore/Reference/Profile/2303652
# Create a temporary file on user's hard drive
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the spatial data file to the temporary file location created above
download.file("https://irma.nps.gov/DataStore/DownloadFile/702822",tmp_geojson)

# Read the temporary spatial data file into an sf object
AKParks = read_sf(tmp_geojson)

# Show a map of where caribou groups were found during the survey. Generate a density heat map also.
#Plot the points over the shapefile (shapefile first, then points)
ggplot() + 
  
  # Park boundary
  geom_sf(data = AKParks,fill = "white", color = "black") + 
  
  # Caribou kernel density
  geom_density_2d(data = data, aes(x = Lon, y = Lat), colour = "red", alpha = 0.9) +
  
  # Caribou observations
  geom_sf(data = spdata, aes(color=factor(Marked)), size= 1, show.legend = T) +
  scale_color_manual(values = c("black","darkgray"))  +
  guides(color=guide_legend(title="Marked (1=True,0=False)")) +
  
  # Adjust colors
  scale_fill_brewer() + # This seems to make the colors less wacky, not sure why
  guides(fill = guide_legend(title = "Density of observations")) + # This gives a legend to the density
  
  # Limit the view port to just WRST
  xlim(-146.5,-139) +
  ylim(59.8,62.65) +
  #ggtitle(paste("Caribou observations distribution and density",sep="") ,subtitle = CompDF$`Start date`) +
  
  # Label axes
  xlab('Longitude') + ylab('Latitude') +
  
  ggtitle('Distribution of Chisana herd during survey') +
  
  # Clean look
  theme_classic()

```

# Results

## Counts

Table X. Counts.
```{r echo=FALSE}

# Raw counts
Sql = paste("SELECT Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull, Adult, Caribou, Unknown, [Total groups observed], [Minimum group size], [Maximum group size], Marked, Seen, [Marked and Seen], [Marked and Not Seen], 
 [Groups in the survey area], [Groups out of the survey area], [Total collared animals observed]
FROM Summary_Census
WHERE (Herd = 'Chisana') AND (Year = ",Year,")",sep="")

data = sqlQuery(Connection,Sql)
data_t = t(data)
colnames(data_t) = 'Counts'
kable(data_t)
```

## Composition

Table X. Composition.
```{r echo=FALSE}

# Raw counts
Sql = paste("SELECT        [Calves/100 cows], [Bulls/100 cows], [Pct. calf], [Pct. cow], [Pct. small bull], [Pct. medium bull], [Pct. large bull], [Pct. small bull (of total bulls)], [Pct. medium bull (of total bulls)], [Pct. large bull (of total bulls)], 
                         [Pct. bull (composition survey)], [Total groups observed], GroupSize_Mean, [Minimum group size], [Maximum group size]
FROM            Summary_Census
WHERE (Herd = 'Chisana') AND (Year = ",Year,")",sep="")

data = sqlQuery(Connection,Sql)
data_t = t(data)
colnames(data_t) = 'Composition'
kable(data_t)
```

## Population Estimate

Note: 2024-06-27 SDMiller: I'm still working out how to run this analysis by trying to reproduce the 2010 Chisana Report put out by Yukon (https://irma.nps.gov/DataStore/DownloadFile/566074)[https://irma.nps.gov/DataStore/DownloadFile/566074]
Example code at J:\Monitoring\Caribou\WRST\Historical Files from Copper Center\Electronic\Chisana\SightabilityModel

Note: Population estimates can only be calculated if sightability information is collected. The next R chunk executes the QC_CaribouGroups_SurveysWithNoSightabilityInformation query to determine if sightability information exists. If sightability information exists you will get a corrected population estimate, otherwise a minimum count.

```{r echo=FALSE}
# We cannot calculate a population estimate based on sightability if no sightability information was collected.
# The two required columns are Marked (whether GPS collared animal(s) were in the group), and Seen, whether the caribou group was observed, or missed.
# If the survey in question shows up in the query then a sightability corrected estimate is not possible
Sql = paste("SELECT  [Herd]
      ,[Year]
      ,[Marked]
      ,[Seen]
      ,[SurveyName]
  FROM [QC_CaribouGroups_SurveysWithNoSightabilityInformation]
WHERE Herd='Chisana' and Year = ",Year,sep="")
data = sqlQuery(Connection,Sql)

```


```{r echo = FALSE}

# If the survey shows up in the list of surveys with no sightability data (above) then we can't calculate a corrected population estimate.
# Show a minimum count
if (nrow(data) > 0){

  # No sightability info was collected for this survey
  print(paste('Sightability information was not collected for the Chisana herd for ',Year,sep=''))
  print('Sightability information appears in the table below')
  kable(data)
  
  print('A minimum count estimate (uncorrected) is possible')
  Sql = paste("SELECT Caribou as [Minimum count total] FROM Summary_Census WHERE Herd='Chisana' and Year = ",Year,"",sep="")
  kable(sqlQuery(Connection,Sql))

}else{
  
  # Calculate a sightability corrected population estimate
  print('Sightability information exists for this survey')
  
  # Observational dataset
  # Note: For some reason Total and Grp_Size are both needed by SightabilityModel R package, even though they are duplicative
  Sql = paste("SELECT  1 as stratum,[Total] as total, 1 as subunit,[Grp_Size] FROM [WRST_Caribou].[dbo].[Sightability_odat] 
  WHERE Herd='Chisana' 
  And Year = ",Year,"
  And [In] = 1 -- This can screw up the pop est if it's all NULLs.
  ",sep="")
  odat = sqlQuery(Connection,Sql)
  
  # Sightability dataset
  Sql = paste("SELECT   ID, Observed, Grp_Size FROM  Sightability_sdat WHERE (Herd = 'Chisana') AND (Year = ",Year,")",sep="")
  sdat = sqlQuery(Connection,Sql)
  
  # Sampling info dataset (essentially all 1s, generate on the fly)
  stratum <- 1
  Nh = 1
  nh = 1
  sampinfo <- data.frame(stratum,Nh,nh)
  
  #Code to fit a sightability model using group size as a covariate
  # Choose the model with the smallest AIC
  fit.group.size <- Sight.Est(form=Observed ~ Grp_Size,sdat=sdat,odat=odat,sampinfo=sampinfo, method="Wong", logCI=T, alpha=0.1, Vm.boot=T, nboot=1000)
  fit.group.size #Provides the results from the model fitting
  print(fit.group.size)
  Summary = summary(fit.group.size)
  Estimate = Summary$tau.hat
  UpperEstimate = Summary$ucl
  LowerEstimate = Summary$lcl
  print(paste("Estimate (group size): ",Estimate," (",LowerEstimate," - ",UpperEstimate,")"))
  
  #Code to fit a sightability model having a constant sightability (assumes 100% sightability?)
   # Choose the model with the smallest AIC
  fit.constant<-Sight.Est(form=Observed ~ 1, sdat=sdat, odat=odat, sampinfo=sampinfo, method="Wong", logCI=T, alpha=0.1, Vm.boot=T, nboot=1000)
  fit.constant  #Provides the results
  print(fit.constant)
  Summary = summary(fit.constant)
  Estimate = Summary$tau.hat
  UpperEstimate = Summary$ucl
  LowerEstimate = Summary$lcl
  print(paste("Estimate (assumes 100% sightability): ",Estimate," (",LowerEstimate," - ",UpperEstimate,")"))
  
  
  #The AICc for the constant model is lower than for the Group Size model so I chose the contant model as most supported.
  # Sight.Est doesn't calculate AICc (the small sample correction for AIC), I did that myself.
  
  #To get the actual sightability correction factor use the following codes
  #This fits the sightability model
  fit.glm <- glm(Observed~1, data=sdat, family=binomial()) 
  
  #This gives you the sightability correction factor (theta)
  theta.SS <- 1+exp(-coef(fit.glm)-summary(fit.glm)$cov.unscaled/2) 
  theta.SS # Provides the actual number
}


```


# Historical Results

```{r echo=FALSE}

# Get the chisana population estimates from the reports to build a chart of population size over time
Sql = paste("SELECT Year
, SourceID
, [Herd (report)]
, FallSurveyDate
, FallEstimatedBulls
, FallPopulationEstimate
, FallPopulationEstimateLower
, FallPopulationEstimateUpper
, FallPopulationEstimateMethod
, FallCows, FallCalves
, FallEstimatedCalves
, FallBull
, FallTotal
, FallCalvesPer100Cows
, FallBullsPer100Cows
, FallPctCow
, FallPctCalves
, FallPctBulls
, FallPctSmallBullsOfBulls
, FallPctMediumBullsOfBulls
, FallPctLargeBullsOfBulls
, FallSampleSize
, ReportYear
, Herd
, Agency
, EarliestContentYear
, LatestContentYear
, Citation
,ResultsValidatedDate
FROM            Summary_ChisanaHistoricalSurveyResults
ORDER BY Year, SourceID
",sep="")
data = sqlQuery(Connection,Sql)

# Source data reports
# Sql = paste("SELECT Citation,SourceID FROM Reports WHERE SourceID in (54,46,47)",sep="")
# sourcesDataFrame = sqlQuery(Connection,Sql)

```


```{r echo=FALSE}

ggplot(
  data %>% filter(FallBullsPer100Cows > 0)
  ) +
  geom_point(aes(x=Year,y=FallCalvesPer100Cows,shape=factor(SourceID))) +
  geom_point(aes(x=Year,y=FallBullsPer100Cows,shape=factor(SourceID))) +
  
  geom_line(aes(x=Year,y=FallBullsPer100Cows,linetype='solid')) +
  geom_line(aes(x=Year,y=FallCalvesPer100Cows,linetype='dashed')) +
  geom_hline(aes(yintercept=35,linetype='dotted')) + # Mgmt goal 35 bulls/100 cows
  geom_hline(aes(yintercept=15),linetype='dotted') + # Mgmt goal 15 calves/100 cows
  
  theme_minimal() +
  ylab("Sex and age ratios (per 100 cows)") +   # Y axis label
  xlab("Year") +# X axis label
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) + # Make X axis labels vertical
  scale_x_continuous(breaks = seq(1978, max(data$Year), by = 2)) + # Set the X axis scale
  
  # Manually adjust points 
  scale_shape_discrete(name="Source", breaks=c("46", "47", "54","999"), labels=c("Cutting, 2023", "Hegel, et al., 2014", "Chisana Caribou Herd\n Working Group, 2012","WRST")) +
  
# Manually adjust lines 
  scale_linetype_discrete(name="Ratio", breaks=c("solid","dashed"), labels=c("Bulls/100 cows","Calves/100 cows"))
  
```

Figure X. Chisana sex and age ratios over time. Dotted horizontal lines show the management goals of 35 bulls/100 cows and 15 calves/100 cows.

```{r echo=FALSE}

ggplot(
  data %>% filter(SourceID == 54 | SourceID == 46 | SourceID == 47) %>% # These are the good, valid reports
  filter(!(SourceID == 54 & Year > 2002))  # Exclude CCH working group estimates after 2002, USGS estimates in Hegel duplicates these estimates and have error bars
  ) +
  geom_point(aes(x=Year,y=FallPopulationEstimate,shape=factor(SourceID))) +
  geom_line(aes(x=Year,y=FallPopulationEstimate)) +
  geom_errorbar(aes(x=Year,ymin=FallPopulationEstimateLower,ymax=FallPopulationEstimateUpper)) +
  
  theme_minimal() +
  ylab("Population Estimate") +   # Y axis label
  xlab("Year") +# X axis label
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) + # Make X axis labels vertical
  scale_y_continuous(breaks = seq(0, 2000, by = 100)) + # Set the X axis scale
  scale_x_continuous(breaks = seq(1978, max(data$Year), by = 2)) + # Set the X axis scale
  #guides(shape=guide_legend(title="Source")) +
  scale_shape_discrete(name="Source",
                         breaks=c("46", "47", "54"),
                         labels=c("Cutting (2023)", "Hegel, et al. (2014)", "Chisana Caribou Herd\n Working Group (2012)"))
  
```

Figure X. Sex and age ratios over time. Captive rearing program was begun in 2003 (CCHWG, 2012).

# References

Chisana Caribou Herd Working Group. 2012. Management plan for the Chisana caribou herd: 2010-2015. Government of Yukon Department of Environment Whitehorse, YT. 48pp.

Cutting, K. 2023. WRANGELL-ST. ELIAS NATIONAL PARK AND PRESERVE WILDLIFE REPORT, Spring, 2023. Unpublished Report. United States Department of the Interior. National Park Service. Copper Center, AK.

HEGEL, T., T. BENTZEN, R. DRUMMOND, J. PUTERA, S. TAYLOR, AND J. WELLS. 2014. Chisana caribou herd population estimate, 2013. Yukon Fish and Wildlife Branch Report TR-14-XX. Whitehorse, Yukon, Canada.
