
```{r echo=FALSE}

# Set report parameters
Year = 2022
Herd = 'Chisana'
SurveyName = '2022 Fall Chisana Caribou Composition Survey' # Change this to the survey name in the WRST Caribou database.
```

---
title: Results from the `r Year` Composition Count Survey of the `r Herd` Caribou Herd in Wrangell - St. Elias National Park and Preserve, Alaska
author: "Scott D. Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(RODBC)
library(sf)
library(ggplot2)

WRST = "Wrangell - St. Elias National Park and Preserve"

# Database connection string
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')

# Function that returns a data frame from an SQL query on the database
# SQL: A valid SQL query
GetDataFrame <- function(SQL) { # create a function with the name my_function
  DF = sqlQuery(Connection,SQL)
  return(DF)
}

# Function to convert a data frame of Lat/Lon pairs to a spatial data frame
# DataFrame: A data frame containing a latitude field and a longitude field
# LatitudeFieldName: The name of the field containing latitude data
# LongitudeFieldName: The name of the field containing longitude data
GetPointSpatialDataFrameFromDataFrame = function(DataFrame,LatitudeFieldName,LongitudeFieldName){
  
  # Convert the data frame into a spatial data frame
  SDF <- sf::st_as_sf(DataFrame, coords = c(LongitudeFieldName,LatitudeFieldName))

  # Set the coordinate system to GCS Lat\Lon WGS1984
  st_crs(SDF) <- st_crs(4326)
  
  return(SDF)
}


TableCount = 1
FigureCount = 1

```

Scott D. Miller  
Information Technology Specialist/Data Manager  
National Park Service, Arctic Inventory and Monitoring Network  
240 W. 5th Ave.  
Anchorage, AK 99501


# Introduction

```{r echo=FALSE}
Sql = paste("SELECT        [Survey name], Herd, [Survey type], [Type of survey], Cow, Calf, [Small bull], [Medium bull], [Large bull], [Bull (composition survey)] AS Bull, [Adult (composition survey)] AS Adult, [Caribou (composition survey)] AS Caribou, 
 Unknown, [Calf, male], [Calf, female], [Calf, unclassified], GroupSize_Mean, [Pct. calves], [Pct. cows], [Pct. bull (composition survey)] AS [Pct. bull], [Calves/100 cows], 
 [Bulls/100 cows (composition survey)] AS [Bulls/100 cows], [Total groups observed], [Groups in the survey area], [Groups out of the survey area], [Marked groups], [Marked groups seen], [Marked groups not seen], 
 [Percent marked groups seen], [Percent marked groups not seen], [Total collared animals observed], [Number of frequencies that were available on survey date (Animal_Movement)], [Collars not heard or not searched], 
 [Total groups marked seen], [Start date], [End date], [Survey days], [Search areas]
FROM            Summary_Census
WHERE ([Survey name] = '",SurveyName,"')",sep="")
CompDF = GetDataFrame(Sql)

# Basic survey details
StartDate = CompDF[1,]$`Start date`
EndDate = CompDF[1,]$`End date`
SurveyDays = CompDF[1,]$`Survey days`

```

```{r echo=FALSE}

Sql = paste("SELECT        Herd, [Survey type], Earliest, [Latest survey], [Years of record], [Survey days], Surveys 
FROM Summary_LengthOfRecordByHerdAndSurveyType Where Herd = '",Herd,"' And [Survey type] = 'Composition count'",sep="")
DF = GetDataFrame(Sql)
if(nrow(DF) > 0){
  Earliest = DF[1,]$Earliest
  TotalSurveys = DF[1,]$Surveys
}
```



Caribou are an integral component of the ecosystem in `r WRST` (WRST) and are of great importance to the ecosystem and to people for both consumptive and non-consumptive uses. WRST is mandated by the Alaska National Interest Lands Conservation Act to protect habitat for caribou, to provide the opportunity for continued subsistence uses, and to conserve natural and healthy wildlife populations. In order to meet these requirements the Park monitors the `r Herd` caribou herd with the goal of providing Park managers with reliable scientific information on the herd's abundance, composition and distribution. 
WRST conducted a herd composition survey from `r StartDate` to `r EndDate` to assess calf recruitment, adult sex ratios, and seasonal distribution. In this report we provide provisional survey results in the form of data and summary statistics, but leave in-depth statistical analysis of population size and composition for a later date.

# Methods

Monitoring of caribou populations in CAKN employs the use of radiocollars and radiotelemetry to locate groups and to provide a mark/recapture estimate of population size and age and sex composition. The use of radiotelemetry is standard throughout Alaska and parts of Canada for monitoring caribou populations.
A sample of 30-40 radiocollared cows is maintained for population assessment. Population assessment is made in two efforts: a post-calving (June) census, when cows are grouped and calf production can be determined; and a fall composition count when bulls associate with cows during the rut. With the mark-recapture estimate of cows from the spring survey and the herd composition obtained from the fall count, the herd size, composition, and calf recruitment can be estimated. Both annual surveys are conducted by aerial observation from a helicopter and small airplane. These caribou monitoring methods were formalized in a collaboration between the WRST and the Central Alaska Inventory and Monitoring Program in 2018 (Putera and Miller, 2018). The WRST caribou monitoring protocol, and standard operating procedures, as well as reports, journal articles and data are available through the NPS's Integrated Resource Management Applications portal at [https://irma.nps.gov/DataStore/Reference/Profile/2217176](https://irma.nps.gov/DataStore/Reference/Profile/2217176). 

#Results

The composition survey took place from `r StartDate` to `r EndDate` over `r SurveyDays` days (Table `r TableCount`). Survey conditions were [fill in survey conditions if available.]. 

Table `r TableCount`. Survey dates and observer/pilot pairs, `r Year`.

```{r echo=FALSE}
Sql = paste("SELECT  DISTINCT CONVERT(Date, SightingDate) AS [Date],   SearchArea, Observer1 AS Observer, Pilot
FROM            CaribouGroups
WHERE        (SurveyName = '",SurveyName,"') AND (SurveyType = 'CC')
ORDER BY [Date],SearchArea, Observer, Pilot",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)
# TableCount = TableCount + 1
```


## Herd distribution

Caribou were distributed as shown in Figure `r FigureCount`.

```{r echo=FALSE}

# Get the spatial data for comp and population surveys
Sql = paste("SELECT SurveyName,SearchArea, GroupNumber, Lat, Lon,Lat as Latitude,Lon as Longitude FROM Dataset_Census WHERE SurveyName = '",SurveyName,"' And Lat > 0",sep="")
CCDF = GetDataFrame(Sql)

# Convert the data into a spatial data frame
# Get rid of rows with no spatial coordinates - they will cause the map to bomb
WithCoordinatesDF = subset(CCDF,Lat != "NULL" & Lon != "NULL")
CCSDF = GetPointSpatialDataFrameFromDataFrame(WithCoordinatesDF,"Lat","Lon")

# Read the shapefile into a spatial features dataset
AKParksShp <- sf::read_sf("C:/Work/GIS Common Layers/GCS/AKParks_reprojected.shp")

#Plot the points over the shapefile (shapefile first, then points)
ggplot() + geom_sf(data = AKParksShp) + 
  geom_sf(data = CCSDF, color='black', size= 2, show.legend = T) +
  # geom_text(data = CCSDF, size = 3, color = 'red',mapping = aes(CCSDF$Longitude,CCSDF$Latitude,label = CCSDF$SurveyName), vjust = -1,hjust = 0,inherit.aes = FALSE) +
  
   # Density layer
    geom_density_2d(data = CCDF, aes(x = Longitude, y = Latitude), colour = "red", alpha = 0.9) +
    scale_fill_brewer() + # This seems to make the colors less wacky, not sure why
    guides(fill = guide_legend(title = "Density of observations")) + # This gives a legend to the density
  
  # Limit the view port to just WRST
  xlim(-146.5,-139) +
  ylim(59.8,62.65) +
  #ggtitle(paste("Composition and population survey observations,",Year,sep="") ,subtitle = "Wrangell-St. Elias National Park and Preserve") +
  xlab('Longitude') + ylab('Latitude')

```

Figure `r FigureCount`. Distribution and density of caribou groups in Wrangell - St. Elias National Park and Preserve and outlying areas during the `r SurveyName`.

```{r echo=FALSE}
FigureCount = FigureCount + 1
```


## Herd Composition

```{r echo=FALSE}
Caribou = CompDF[1,]$Caribou
BullsTo100Cows = CompDF[1,]$`Bulls/100 cows`
BullsTo100Cows = round(BullsTo100Cows, digits = 1) 
CalvesTo100Cows = CompDF[1,]$`Calves/100 cows`
CalvesTo100Cows = round(CalvesTo100Cows, digits = 1) 

```

We classified `r Caribou` caribou during the survey. Calf recruitment was `r CalvesTo100Cows` per 100 cows and there were `r BullsTo100Cows` bulls per 100 cows (Table `r TableCount`).

Table `r TableCount`. Composition survey results.
```{r echo=FALSE}
kable(t(CompDF))
TableCount = TableCount + 1

```

## Historical Results

### Counts

```{r echo=FALSE}
Sql = paste("SELECT [Year],Season,  Cow, Calf, [Small bull], [Medium bull], [Large bull], [Bull (composition survey)] AS Bull, [Adult (composition survey)] AS Adult, [Caribou (composition survey)] AS Caribou
,[Total groups observed] as n
FROM            Summary_Census
WHERE Herd = '",Herd,"' And [Survey type] = 'CC' 
ORDER BY Year DESC,Season DESC",sep="")
ComparisonDF = GetDataFrame(Sql)

kable(ComparisonDF)

```

## Composition

```{r echo=FALSE}
Sql = paste("SELECT Year, Season
, GroupSize_Mean as [Mean group size], [Pct. calves], [Pct. cows], [Pct. bull (composition survey)] AS [Pct. bull], [Calves/100 cows], 
 [Bulls/100 cows (composition survey)] AS [Bulls/100 cows], [Total groups observed] as n
FROM            Summary_Census
WHERE Herd = '",Herd,"' And [Survey type] = 'CC' 
ORDER BY Year DESC",sep="")
ComparisonDF = GetDataFrame(Sql)

kable(ComparisonDF)

ggplot(data=ComparisonDF, aes(x=Year, y=`Calves/100 cows`, group=1)) +
  geom_line()+
  geom_point()

#plot(ComparisonDF$`Calves/100 cows` ~ ComparisonDF$Year)

```

Figure `r FigureCount`. Calves/100 cows over time.
```{r echo=FALSE}
ggplot(data=ComparisonDF, aes(x=Year, y=`Calves/100 cows`, group=1)) +
  geom_line()+
  geom_point()

#plot(ComparisonDF$`Calves/100 cows` ~ ComparisonDF$Year)

FigureCount = FigureCount + 1
```



# Sightability

Stats

```{r echo=FALSE}
Sql = paste("SELECT Year,Season,[Total groups observed] as n
-- , [Groups in the survey area], [Groups out of the survey area] -- the [In] attribute isn't used much
, [Marked groups], [Marked groups seen]
--, [Marked groups not seen]
, [Percent marked groups seen]
--, [Percent marked groups not seen]
--, [Total collared animals observed], [Number of frequencies that were available on survey date (Animal_Movement)]
, [Collars not heard or not searched], 
 [Total groups marked seen]
FROM            Summary_Census
WHERE Herd = '",Herd,"' And [Survey type] = 'CC' 
ORDER BY Year DESC",sep="")
ComparisonDF = GetDataFrame(Sql)

kable(ComparisonDF)

```


<!-- Table `r TableCount`. Composition survey results by search area. [DEFINE THE SEARCH AREAS]. -->
```{r echo=FALSE}
# Sql = paste("SELECT        [Survey name], Herd, [Type of survey], SearchArea, Cow, Calf, [Small bull], [Medium bull], [Large bull], [Bull (composition survey)], [Adult (composition survey)], [Caribou (composition survey)], Unknown, [Calf, male], 
#  [Calf, female], [Calf, unclassified], [Calves/100 cows], [Bulls/100 cows (composition survey)], [Total groups observed], [Groups in the survey area], [Groups out of the survey area], [Marked groups], [Marked groups seen], 
#  [Marked groups not seen], [Percent marked groups seen], [Percent marked groups not seen], [Total collared animals observed], [Total groups marked seen], [Start date], [End date], [Survey days], [Search areas], Park, Year
# FROM Summary_CensusBySearchArea
# WHERE ([Survey name] = '",SurveyName,"')
# ORDER BY [Searcharea]",sep="")
# CCDF = GetDataFrame(Sql)
# kable(t(CCDF))
# TableCount = TableCount + 1
```

# Discussion


# References
ANILCA, 1980

# Appendix A: Dataset Certification Summary

Appendix A contains a summary of data quality certification levels and number of records for the `r SurveyName`. Definitions follow the summary table.

```{r echo=FALSE}
Sql = paste("SELECT SurveyName, CertificationLevel, [Number of records]
FROM     Summary_Census_CertificationLevel
WHERE ([SurveyName] = '",SurveyName,"')",sep="")
DF = GetDataFrame(Sql)
kable(DF)
```

CertificationLevel | Description
--- | ---
Raw | Records imported directly from field surveys with no quality control or validation procedures performed.
Provisional | Records that have undergone minimal quality control and/or have been examined by the project leader and are deemed suitable for internal reports but have not been validated against a published summary or report. Some records may never be fully certified due to the lack of validating materials.
Certified | Records that have undergone extensive quality control procedures and have been validated against published reports, journal articles or data summary reports.



# Appendix B: Survey Description and Notes 

Appendix B contains a description of the survey and any notes about the survey as written by the pilot, observer or biologist. These data are available in the WRST_Caribou.Surveys table.

```{r echo=FALSE}
# Survey description and notes are stored in the Surveys table. See if there is a record for the survey and store 
# the description and notes in variables.
Sql = paste("SELECT SurveyName, Description, Notes FROM Surveys WHERE ([SurveyName] = '",SurveyName,"')",sep="")
DF = GetDataFrame(Sql)
Description = ""
SurveyNotes = ""
if(nrow(DF) > 0){
  Description = DF[1,]$Description
  SurveyNotes = DF[1,]$Notes
}
```

## Survey Description
`r Description`

## Survey Notes
`r SurveyNotes`

# Appendix C: Data Quality Issues

Appendix C Contains known data quality issues that potential users of the dataset should know. 

```{r echo=FALSE}
# Survey description and notes are stored in the Surveys table. See if there is a record for the survey and store 
# the description and notes in variables.
Sql = paste("SELECT Surveys_Issues.IsResolved, Surveys_Issues.DateLogged, Surveys_Issues.Author, Surveys_Issues.Issue
FROM  Surveys INNER JOIN
 Surveys_Issues ON Surveys.SurveyName = Surveys_Issues.SurveyName where Surveys.SurveyName = '",SurveyName,"'
ORDER BY Surveys.SurveyName DESC,Surveys_Issues.IsResolved, Surveys_Issues.DateLogged",sep="")
DF = GetDataFrame(Sql)
kable(DF)

```


# Appendix D: Quality Control Checks Results

Appendix D contains the results of quality control checks on the survey dataset. These results are intended to inform NPS scientists about any data quality problems and should be resolved, if possible, before publication.Move unresolvable issues that data consumers should know about to the Surveys_Issues table and publish from there.

## QC-1: Observed/Available Collars Reconciliation
Ideally, every collar that is registered in the NPS Animal Movement database for the `r Herd` monitoring program is searched for during a composition survey. Table `r TableCount` shows the count of frequencies detected during the `r SurveyName` and the number of collars that were available according to the data in the Animal Movement database. It is not critical that the numbers match, only that they reconcile in a logical way. Collars occasionally fail, the terrain may block their signal, the crew ran out of time or encountered bad weather, or collars were are not detected for another reason. Sometimes collar deployments are not terminated in Animal_Movement after a mortality, resulting in an inflated inventory of available collars. The list below is a starting point for reconciling these issues.

Table `r TableCount`. Observed/Available Collars Reconciliation.

```{r echo=FALSE}
Sql = paste("SELECT 
-- Herd, [Survey type], 
SurveyName as [Survey], [Count of frequencies] as [Observed collar frequencies], AvailableCollars as [Available collar frequencies]
FROM     QC_CaribouGroups_FrequenciesCountNotMatchingAvailableFrequenciesInAnimal_Movement WHERE (SurveyName = '",SurveyName,"')",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)

TableCount = TableCount + 1
```

## QC-2: Marked groups/frequency(ies) detected check
An important component of caribou surveys is sightability which is accomplished through a mark-recapture model. Groups are marked via GPS collars. The first survey flight does not have access to GPS telemetry from the collars and searches for caribou blind. The follow up flight is able to find marked groups via telemetry and tags each observed animal group as marked if it had at least one collared animal in it and the group is marked as 'seen' if the blind group found it. This quality control test shows records having a non-null FrequenciesInGroup attribute and its corresponding Marked attribute marked anything but 1.

Table `r TableCount`. Records having non-null FrequenciesInGroup attribute with Marked attribute not equal to 1 (boolean, True). A blank table below indicates a QC check PASS.

```{r echo=FALSE}
Sql = paste("SELECT SurveyName, SightingDate, GroupNumber, Marked, FrequenciesInGroup
FROM QC_CaribouGroups_GroupsWithFrequenciesDetectedButMarkedIsNULL
WHERE (SurveyName = '",SurveyName,"')",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)

TableCount = TableCount + 1
```

## QC-3: Missing sightability information
Quality control check QC-3 extends check QC-2 by providing information about how many survey records have null Seen and Marked attributes. 

Table `r TableCount`. Surveys where sightability data (completely null Seen and Marked attributes for all survey records) were not collected. A blank table below indicates a QC check PASS.

```{r echo=FALSE}
Sql = paste("SELECT SurveyName as Survey, Seen, Marked
FROM            QC_CaribouGroups_SurveysWithNoSightabilityInformation
WHERE (SurveyName = '",SurveyName,"') ORDER BY SurveyName DESC",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)

TableCount = TableCount + 1
```

## QC-4:
Quality control check QC-3 extends check QC-2 by providing information about how many survey records have null Seen and Marked attributes. 

Table `r TableCount`. Surveys where sightability data (completely null Seen and Marked attributes for all survey records) were not collected. A blank table below indicates a QC check PASS.

```{r echo=FALSE}
Sql = paste("SELECT SurveyName as Survey, Seen, Marked
FROM            QC_CaribouGroups_SurveysWithNoSightabilityInformation
WHERE (SurveyName = '",SurveyName,"') ORDER BY SurveyName DESC",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)

TableCount = TableCount + 1
```




