```{r echo=FALSE}

# Set report parameters
Year = 2023 # Year of the comp count survey
Herd = 'Chisana' # Herd
#SurveyName = paste(Year," Fall ",Herd," Caribou Composition Survey",sep="") # Change this to the survey name in the WRST Caribou database.
SurveyName = "2007-10-13 Chisana Caribou Population Survey"


```

---
title: "Results from the `r Year` Composition Count Survey of the `r Herd` Caribou
  Herd in Wrangell - St. Elias National Park and Preserve, Alaska"
author: ""
date: ""
output:
  html_document:
  toc: yes
---

```{r setup, include=FALSE}

# Suppress warnings
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Setup
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(RODBC)
library(sf)
library(ggplot2)
library(tidyverse)

# Restrain the number of decimal places
options(digits = 1) 

# WRST Shortcut
WRST = "Wrangell - St. Elias National Park and Preserve"

# Database connection string
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')

# Function that returns a data frame from an SQL query on the database
# SQL: A valid SQL query
GetDataFrame <- function(SQL) { # create a function with the name my_function
  DF = sqlQuery(Connection,SQL)
  return(DF)
}

# Function to convert a data frame of Lat/Lon pairs to a spatial data frame
# DataFrame: A data frame containing a latitude field and a longitude field
# LatitudeFieldName: The name of the field containing latitude data
# LongitudeFieldName: The name of the field containing longitude data
GetPointSpatialDataFrameFromDataFrame = function(DataFrame,LatitudeFieldName,LongitudeFieldName){
  
  # Convert the data frame into a spatial data frame
  SDF <- sf::st_as_sf(DataFrame, coords = c(LongitudeFieldName,LatitudeFieldName))

  # Set the coordinate system to GCS Lat\Lon WGS1984
  st_crs(SDF) <- st_crs(4326)
  
  return(SDF)
}


# Counters to increment the table and figure numbers as the report progresses.
TableCount = 1
FigureCount = 1

```

Scott D. Miller\
Information Technology Specialist/Data Manager\
National Park Service, Arctic Inventory and Monitoring Network\
240 W. 5th Ave.\
Anchorage, AK 99501

Kyle Cutting\
Wildlife Biologist\
National Park Service, `r WRST`\
8 Richardson Hwy Mile 106\
Copper Center, AK 99573

# Introduction

```{r echo=FALSE}

# Get the summarized data for the comp count survey into a data frame
Sql = paste("SELECT        [Survey name], Herd, [Survey type], [Type of survey], Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull, Adult, Caribou, Unknown, [Calf, male], [Calf, female], [Calf, unclassified], Yearling, GroupSize_Mean, [Pct. calf], 
                         [Pct. cow], [Pct. small bull], [Pct. medium bull], [Pct. large bull], [Pct. small bull (of total bulls)], [Pct. medium bull (of total bulls)], [Pct. large bull (of total bulls)], [Pct. bull (composition survey)], [Pct. bull (population survey)], 
                         [Calves/100 cows], [Bulls/100 cows], [Total groups observed], [Marked groups], [Observed groups], [Marked groups that were observed], [Marked groups that were not observed], [Pct. marked groups observed], 
                         [Groups in the survey area], [Groups out of the survey area], [Marked groups in survey area], [Observed groups in survey area], [Marked groups observed in survey area], [Pct. marked groups observed in survey area], 
                         [Total collared animals observed], [Number of frequencies that were available on survey date (Animal_Movement)], [Collars not heard or not searched], [Start date], [End date], [Survey days], [Search areas], Park, Year, 
                         CertificationLevel, DateValidated, ValidatedBy
FROM            Summary_Census
WHERE ([Survey name] = '",SurveyName,"')",sep="")
CompDF = GetDataFrame(Sql)

# Basic survey details
StartDate = CompDF[1,]$`Start date`
EndDate = CompDF[1,]$`End date`
SurveyDays = CompDF[1,]$`Survey days`

```

```{r echo=FALSE}
# Get the historical length of the various surveys into a data frame
Sql = paste("SELECT        Herd, [Survey type], Earliest, [Latest survey], [Years of record], [Survey days], Surveys 
FROM Summary_LengthOfRecordByHerdAndSurveyType Where Herd = '",Herd,"' And [Survey type] = 'Composition count'",sep="")
DF = GetDataFrame(Sql)
if(nrow(DF) > 0){
  Earliest = DF[1,]$Earliest
  TotalSurveys = DF[1,]$Surveys
}
```

Caribou are an integral component of the ecosystem in `r WRST` (WRST) and are of great importance to the ecosystem and to people for both consumptive and non-consumptive uses. WRST is mandated by the Alaska National Interest Lands Conservation Act to protect habitat for caribou, to provide the opportunity for continued subsistence uses, and to conserve natural and healthy wildlife populations. In order to meet these requirements the Park monitors the `r Herd` caribou herd with the goal of providing Park managers with reliable scientific information on the herd's abundance, composition and distribution. To further this mission, WRST conducted a herd composition survey from `r StartDate` to `r EndDate` to assess calf recruitment, adult sex ratios, and seasonal distribution. This report summarizes the results of this survey.

# Methods

WRST conducts aerial caribou surveys using radiocollars and radiotelemetry to locate animal groups and to provide a mark/recapture estimate of population size and age and sex composition. The use of radiotelemetry is standard throughout Alaska and parts of Canada for monitoring caribou populations. A sample of 30-40 radiocollared cows is maintained for population assessment. Population assessment is made in two efforts: a post-calving (June) census, when cows are grouped and calf production can be determined; and a fall composition count when bulls associate with cows during the rut. With the mark-recapture estimate of cows from the spring survey and the herd composition obtained from the fall count, the herd size, composition, and calf recruitment can be estimated. Both annual surveys are conducted using a helicopter and small airplane. These caribou monitoring methods were formalized in a collaboration between the WRST and the Central Alaska Inventory and Monitoring Program in 2018 (Putera and Miller, 2018). The WRST caribou monitoring protocol, and standard operating procedures, as well as reports, journal articles and data are available through the NPS's Integrated Resource Management Applications portal at <https://irma.nps.gov/DataStore/Reference/Profile/2217176>.

# Results

The composition survey took place from `r StartDate` to `r EndDate` over `r SurveyDays` days (Table `r TableCount`).

Table `r TableCount`. Survey dates and observer/pilot pairs, `r Year`.

```{r echo=FALSE}
# Get the survey days and observers into a frame
Sql = paste("SELECT  DISTINCT CONVERT(Date, SightingDate) AS [Date],   SearchArea, Observer1 AS Observer, Pilot
FROM            CaribouGroups
WHERE        (SurveyName = '",SurveyName,"')
ORDER BY [Date],SearchArea, Observer, Pilot",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)
```

## Herd distribution

Caribou were distributed as shown in Figure `r FigureCount`.

```{r echo=FALSE}

# Get the spatial data for comp and population surveys
Sql = paste("SELECT SurveyName,SearchArea, GroupNumber, Lat, Lon,Lat as Latitude,Lon as Longitude FROM Dataset_Census WHERE SurveyName = '",SurveyName,"' And Lat > 0",sep="")
CCDF = GetDataFrame(Sql)

# Convert the data into a spatial data frame
# Get rid of rows with no spatial coordinates - they will cause the map to bomb
WithCoordinatesDF = subset(CCDF,Lat != "NULL" & Lon != "NULL")
CCSDF = GetPointSpatialDataFrameFromDataFrame(WithCoordinatesDF,"Lat","Lon")


# Park boundaries (geojson) are at https://irma.nps.gov/DataStore/Reference/Profile/2303652
# Create a temporary file on user's hard drive
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the spatial data file to the temporary file location created above
download.file("https://irma.nps.gov/DataStore/DownloadFile/702822",tmp_geojson)

# Read the temporary spatial data file into an sf object
AKParks = read_sf(tmp_geojson)

# Show a map of where caribou groups were found during the survey. Generate a density heat map also.
#Plot the points over the shapefile (shapefile first, then points)
ggplot() + 
  
  # Park boundary
  geom_sf(data = AKParks,fill = "white", color = "black") + 
  
  # Caribou kernes density
  geom_density_2d(data = CCDF, aes(x = Longitude, y = Latitude), colour = "red", alpha = 0.9) +
  
  # Caribou observations
  geom_sf(data = CCSDF, color='black', size= 1, show.legend = T) +
  
  # Adjust colors
  scale_fill_brewer() + # This seems to make the colors less wacky, not sure why
  guides(fill = guide_legend(title = "Density of observations")) + # This gives a legend to the density
  
  # Limit the view port to just WRST
  xlim(-146.5,-139) +
  ylim(59.8,62.65) +
  #ggtitle(paste("Caribou observations distribution and density",sep="") ,subtitle = CompDF$`Start date`) +
  
  # Label axes
  xlab('Longitude') + ylab('Latitude') +
  
  # Clean look
  theme_classic()

```

Figure `r FigureCount`. Distribution and density of caribou groups in Wrangell - St. Elias National Park and Preserve and outlying areas during the `r SurveyName`.

```{r echo=FALSE}
# Increment the figure counter
FigureCount = FigureCount + 1
```

## Herd Composition

```{r echo=FALSE}

# Store some key values to be put in the report text further on
Caribou = CompDF[1,]$Caribou
BullsTo100Cows = CompDF[1,]$`Bulls/100 cows`
BullsTo100Cows = round(BullsTo100Cows, digits = 1) 
CalvesTo100Cows = CompDF[1,]$`Calves/100 cows`
CalvesTo100Cows = round(CalvesTo100Cows, digits = 1) 

```

We classified `r Caribou` caribou during the survey. Calf recruitment was `r CalvesTo100Cows` per 100 cows and there were `r BullsTo100Cows` bulls per 100 cows (Tables `r TableCount` and 2).

Table `r TableCount`. Animal counts by sex, size summary, `r SurveyName`.

```{r echo=FALSE}
CountsSql = paste("SELECT
--[Start date]
--, [End date]
--[Survey days]
--, [Search areas]
Cow
, Calf
, [Small bull]
, [Medium bull]
, [Large bull]
, Bull, Adult
, Caribou
, Unknown
--, [Calf, unclassified]
--, [Calf, male], 
--[Calf, female]
--, Yearling
, [Total groups observed]
FROM Summary_Census
WHERE        ([Survey name] = '",SurveyName,"')",sep="")
CountsDF = GetDataFrame(CountsSql)
kable(t(CountsDF))

# Increment the table counter
TableCount = TableCount + 1

```

Table `r TableCount`. Sex and age ratios, and summary statistics.

```{r echo=FALSE}
CountsSql = paste("SELECT GroupSize_Mean
,[Minimum group size]
,[Maximum group size]
, [Calves/100 cows]
, [Bulls/100 cows]
, [Pct. calf]
, [Pct. cow]
, [Pct. small bull]
, [Pct. medium bull]
, [Pct. large bull]
, [Pct. small bull (of total bulls)]
, [Pct. medium bull (of total bulls)]
, [Pct. large bull (of total bulls)]
, [Pct. bull (composition survey)]
, [Pct. bull (population survey)]
--, [Total groups observed] -- When the table is transposed using kable this integer is converted to decimal which looks weird
FROM Summary_Census
WHERE        ([Survey name] = '",SurveyName,"')",sep="")
CountsDF = GetDataFrame(CountsSql)
kable(t(CountsDF))

# Increment the table counter
TableCount = TableCount + 1

```


```{r echo=FALSE}

      #Proportion of total values composition bar chart
Calf = CompDF$Calf / CompDF$Caribou
Cow = CompDF$Cow / CompDF$Caribou
Bull = CompDF$Bull / CompDF$Caribou
SmallBull = CompDF$`Small bull` / CompDF$Caribou
MediumBull = CompDF$`Medium bull` / CompDF$Caribou
LargeBull = CompDF$`Large bull` / CompDF$Caribou
Unknown = CompDF$Unknown / CompDF$Caribou

Values <- c(Calf,Cow,SmallBull,MediumBull,LargeBull,Bull,Unknown)    
Values[is.na(Values)] = 0 # NAs to 0
Categories <- c('Calf','Cow','Sm. Bull','Med. Bull','Lg. Bull','Bull','Unclassified') 
barplot(Values,names.arg = Categories,ylim=c(0,1),main="Herd composition",ylab="Proportion of total")


```

Figure `r FigureCount`. Herd composition by age and sex, proportion of total caribou observed (`r CompDF$Caribou`).

```{r echo=FALSE}
FigureCount = FigureCount + 1
```

## Radiotelemetry 

```{r echo=FALSE}
RTSql = paste("SELECT  
[Total collared animals observed] as [Frequencies detected]
,[Collars not heard or not searched] as [Collars unheard]
, [Number of frequencies that were available on survey date (Animal_Movement)] as [Available collars]
FROM Summary_Census
WHERE ([Survey name] = '",SurveyName,"')",sep="")
RTDF = GetDataFrame(RTSql)
AvailableCollars = RTDF$`Available collars`
FrequenciesDetected = RTDF$`Frequencies detected`
Unheard = RTDF$`Collars unheard`
```

There were `r AvailableCollars` collared caribou in the `r Herd` during this survey, of which `r FrequenciesDetected` were detected. Of the available collars, `r Unheard` were undetected. These missing collars may have been out of the survey area, failed in the field or not searched for because of time, weather or logistical constraints.

Table `r TableCount`. Radiotelemetry summary.

```{r echo=FALSE}
#kable(t(RTDF))
kable(RTDF,col.names = NA,row.names = FALSE,caption='Table 33. Radiotelemetry summary.')
# Increment the table counter
TableCount = TableCount + 1

```


## Sightability Metrics

Table `r TableCount`. Sightability results (sightability results are more pertinent to population surveys than composition surveys, but are included here for completeness).

```{r echo=FALSE}
SightabilitySql = paste("SELECT 
'--' as [In search area]
, [Groups in the survey area]
, [Marked groups in survey area] -- In = 1 And Marked = 1
, [Marked groups observed in survey area] -- In = 1 And Marked  = 1
, Convert(Varchar(1000),Format([Pct. marked groups observed in survey area],'##')) + '%' as [Pct. marked groups observed in survey area]
--, [Observed groups in survey area] -- In = 1 And Seen = 1

, '--' as [Out of search area]
, [Groups out of the survey area]

, '--' as [All groups]
, [Total groups observed]
, [Marked groups]
, [Observed groups]
, [Marked groups that were observed]
, [Marked groups that were not observed]
, iif([Pct. marked groups observed] > 0,Convert(Varchar(1000),Format([Pct. marked groups observed],'##')) + '%',NULL)  as [Pct. marked groups observed]
, [Total collared animals observed]
FROM            Summary_Census
WHERE        ([Survey name] = '",SurveyName,"')",sep="")
SightabilityDF = GetDataFrame(SightabilitySql)

kable(t(SightabilityDF))

# Increment the table counter
TableCount = TableCount + 1

```


## Historical Results

Tables `r TableCount` and `r TableCount + 1` provide the most recent survey data compared to the results of past composition count surveys for the `r Herd` herd.

### Counts

Table `r TableCount`. Historical composition of the `r Herd` herd.

```{r echo=FALSE}

# Grab all the historical comp count data for the herd into a frame and show it
# Exclude raw data - provisional or certified only
Sql = paste("SELECT [Year], [Survey type], Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull, Adult, Caribou, Unknown, [Total groups observed] as n
FROM            Summary_Census where CertificationLevel = 'Certified'
And Herd = '",Herd,"' 
And [Survey type] = 'CC' 
And CertificationLevel = 'Certified'
And Year <= ",Year,"
ORDER BY Year",sep="")
ComparisonDF = GetDataFrame(Sql)
kable(ComparisonDF)
TableCount = TableCount + 1
```

Survey types: CC = Composition count, PE = Population estimate

### Composition

Table `r TableCount` and Figures `r FigureCount` and `r FigureCount + 1` summarize historical sex and age ratios and herd composition statistics over time for the `r Herd` herd.

#### Historical Data

Table `r TableCount`. Historical sex and age ratios and herd composition statistict for the `r Herd` herd.

```{r echo=FALSE}
# Grab all the historical comp count statistical data for the herd into a frame and show it (the table above shows counts; this table shows calculated stats)
# Exclude raw data - provisional or certified only
Sql = paste("SELECT Herd, Year,Timing, [Survey type], GroupSize_Mean AS [Mean group size], [Pct. calf], [Pct. cow], [Pct. bull (composition survey)] AS [Pct. bull], [Calves/100 cows], [Bulls/100 cows], [Total groups observed] AS n, [Pct. small bull], 
 [Pct. medium bull], [Pct. large bull]
FROM Summary_Census
WHERE Herd = '",Herd,"' And [Survey type] = 'CC'
And CertificationLevel = 'Certified'
And Year <= ",Year,"
ORDER BY Herd,Year DESC,Timing DESC",sep="")
StatsDF = GetDataFrame(Sql)

kable(StatsDF)

TableCount = TableCount + 1
```
#### Sex and Age Ratios

```{r echo=FALSE}
GraphTitle = paste("Historical sex and age ratios, ",Herd," caribou herd",sep="")

# Plot the sex and age ratios
ggplot() +
  
  # Bulls/100 cows
  geom_line(data=StatsDF %>% filter(is.na(`Bulls/100 cows`)==FALSE),aes(x=Year,y = `Bulls/100 cows`, colour="Bulls/100 cows"),size=1, linetype='solid') +
  
  # Calves/100 cows
  geom_line(data=StatsDF %>% filter(is.na(`Calves/100 cows`)==FALSE),aes(x=Year,y = `Calves/100 cows`,colour="Calves/100 cows"),size=1, linetype='solid') +

  xlab("Year") +
  ylab("Ratio (category of caribou:100 cows)") +
  #ggtitle(GraphTitle) +
  #scale_color_manual(labels = c("Bulls:100 cows", "Calves:100 cows"), values = c("black", "darkgray"))  +
  theme_classic() +
  scale_colour_manual(name="Legend",  values=c(`Bulls/100 cows`="black",`Calves/100 cows`="darkgray"))


FigureCount = FigureCount + 1
```

Figure `r FigureCount`. Historical sex and age ratios,  `r Herd` caribou herd.


#### Herd Composition

```{r echo=FALSE}

# Plot the absolute numbers
GraphTitle = paste("Historical sex and age composition as a proportion of total, ", Herd," caribou herd.",sep="")


ggplot(data=StatsDF %>% filter(Timing == 'Late') %>% arrange(Year,Timing)) + # Subset to get rid of the spring surveys that have 100% calves
  geom_ribbon(aes(x=Year, ymin=0, ymax=`Pct. calf`, fill="Calf")) +
  geom_ribbon(aes(x=Year, ymin=`Pct. calf`, ymax=`Pct. calf` + `Pct. bull`, fill="Bull")) +
  geom_ribbon(aes(x=Year,ymin=`Pct. calf` + `Pct. bull`,ymax=`Pct. calf` + `Pct. bull`+`Pct. cow` ,fill="Cow")) +
  scale_fill_manual("Legend",values=c("darkgray","black","lightgray")) +
  theme_classic()

FigureCount = FigureCount + 1
```

Figure `r FigureCount`. `r GraphTitle`.

# Discussion

[Add discussion here]

# References

ANILCA, 1980

# Appendix A: Dataset Certification Summary

Appendix A contains a summary of data quality certification levels and number of records for the `r SurveyName` (current as of `r Sys.Date()`). Definitions follow the summary table. Data sources appear after the certification summary.

```{r echo=FALSE}

# This shows how trustworthy the dataset is. See the table of definitions following
Sql = paste("SELECT SurveyName, CertificationLevel, [Number of records]
FROM     Summary_Census_CertificationLevel
WHERE ([SurveyName] = '",SurveyName,"')",sep="")
DF = GetDataFrame(Sql)
kable(DF)
```

| CertificationLevel | Description                                                                                                                                                                                                                                                                                            |
|--------------------|----------------------------------------------------|
| Raw                | Records imported directly from field surveys with no quality control or validation procedures performed.                                                                                                                                                                                               |
| Provisional        | Records that have undergone minimal quality control and/or have been examined by the project leader and are deemed suitable for internal reports but have not been validated against a published summary or report. Some records may never be fully certified due to the lack of validating materials. |
| Certified          | Records that have undergone extensive quality control procedures and have been validated against published reports, journal articles or data summary reports.                                                                                                                                          |

Data was imported into the database from the following sources:

Table `r TableCount`. Survey data source(s).

```{r echo=FALSE}
# A description of the survey and/or notes provided by personnel is stored in the database Surveys table.
# These data may be useful for someone to know - add it in this appendix
# Survey description and notes are stored in the Surveys table. See if there is a record for the survey and store 
# the description and notes in variables.
Sql = paste("SELECT DISTINCT SourceFilename AS [Source file]
, '[' + SourceFileURL + '](' + SourceFileURL + ')' As SourceFileURL, RecordInsertedBy
,Convert(Date,RecordInsertedDate)
FROM            CaribouGroups
WHERE ([SurveyName] = '",SurveyName,"')",sep="")
DF = GetDataFrame(Sql)
kable(DF)
TableCount = TableCount + 1
```

# Appendix B: Survey Description and Notes

Appendix B contains a description of the survey and any notes about the survey as written by the pilot, observer or biologist. These data are available in the WRST_Caribou.Surveys table.

```{r echo=FALSE}
# A description of the survey and/or notes provided by personnel is stored in the database Surveys table.
# These data may be useful for someone to know - add it in this appendix
# Survey description and notes are stored in the Surveys table. See if there is a record for the survey and store 
# the description and notes in variables.
Sql = paste("SELECT SurveyName, Description, Notes FROM Surveys WHERE ([SurveyName] = '",SurveyName,"')",sep="")
DF = GetDataFrame(Sql)
Description = ""
SurveyNotes = ""
if(nrow(DF) > 0){
  Description = DF[1,]$Description
  SurveyNotes = DF[1,]$Notes
}
```

## Survey Description

`r Description`

## Survey Notes

`r SurveyNotes`

# Appendix C: Data Quality Issues

Appendix C Contains known data quality issues that potential users of the dataset should know (current as of `r Sys.Date()`).

```{r echo=FALSE}

# Data quality issues are inevitable and stored in the Surveys_Issues database table. Dump them out so the user can see what they are dealing with.
Sql = paste("SELECT Surveys_Issues.IsResolved, Surveys_Issues.DateLogged, Surveys_Issues.Author, Surveys_Issues.Issue
FROM  Surveys INNER JOIN
 Surveys_Issues ON Surveys.SurveyName = Surveys_Issues.SurveyName where Surveys.SurveyName = '",SurveyName,"'
ORDER BY Surveys.SurveyName DESC,Surveys_Issues.IsResolved, Surveys_Issues.DateLogged",sep="")
DF = GetDataFrame(Sql)
kable(DF)

```

# Appendix D: Quality Control Checks Results

Current as of `r Sys.Date()`

Appendix D contains the results of quality control checks on the survey dataset. These results are intended to inform NPS scientists about any data quality problems and should be resolved, if possible, before publication.

## QC-1: Observed/Available Collars Reconciliation

Ideally, every collar that is registered in the NPS Animal Movement database for the `r Herd` monitoring program is searched for during a composition survey. Table `r TableCount` shows the count of frequencies detected during the `r SurveyName` and the number of collars that were available according to the data in the Animal Movement database. It is not critical that the numbers match, only that they reconcile in a logical way. Collars occasionally fail, the terrain may block their signal, the crew ran out of time or encountered bad weather, or collars were are not detected for another reason. Sometimes collar deployments are not terminated in Animal_Movement after a mortality, resulting in an inflated inventory of available collars. The list below is a starting point for reconciling these issues.

Table `r TableCount`. Observed/Available Collars Reconciliation.

```{r echo=FALSE}
Sql = paste("SELECT 
--[Herd],[Survey type],[SurveyName]
[Count of frequencies] as [Frequencies detected]
      ,[AvailableCollars] as [Collars that were available (Animal_Movement)]
      --,[Year]
  FROM [WRST_Caribou].[dbo].[QC_CaribouGroups_SurveysHavingFrequenciesCountNotMatchingAvailableFrequenciesInAnimal_Movement] WHERE (SurveyName = '",SurveyName,"')",sep="")
QCDF = GetDataFrame(Sql)
kable(QCDF)

TableCount = TableCount + 1
```

## QC-2: Marked groups/frequency(ies) detected check

An important component of caribou surveys is sightability which is accomplished through a mark-recapture model. Groups are marked via GPS collars. The first survey flight does not have access to GPS telemetry from the collars and searches for caribou blind. The follow up flight is able to find marked groups via telemetry and tags each observed animal group as marked if it had at least one collared animal in it and the group is marked as 'seen' if the blind group found it. This quality control test shows records having a non-null FrequenciesInGroup attribute and its corresponding Marked attribute marked anything but 1.

Table `r TableCount`. Records having non-null FrequenciesInGroup attribute with Marked attribute not equal to 1 (boolean, True). A blank table below indicates a QC check PASS.

```{r echo=FALSE}
Sql = paste("SELECT SurveyName, SightingDate, GroupNumber, Marked, FrequenciesInGroup
FROM QC_CaribouGroups_GroupsWithFrequenciesDetectedButMarkedIsNULL
WHERE (SurveyName = '",SurveyName,"')",sep="")
QCDF = GetDataFrame(Sql)
kable(QCDF)

TableCount = TableCount + 1
```

## QC-3: Missing spatial coordinates

Quality control check QC-3 returns the number of survey records having NULL spatial coordinates or spatial coordinates equal to zero.

Table `r TableCount`. Bad spatial coordinates. A blank table below indicates a QC check PASS.

```{r echo=FALSE}
Sql = paste("SELECT *
FROM QC_CaribouGroups_SurveysMissingSpatialCoordinates WHERE (SurveyName = '",SurveyName,"')",sep="")
QCDF = GetDataFrame(Sql)
kable(QCDF)

TableCount = TableCount + 1
```

## QC-5: Historical counts of caribou by sex and age

**NOTE: The chart below includes uncertified data and is meant only for quality control purposes.**

Analytically, plotting absolute counts of caribou has little value (you would standardize by proportion of the total by year, see above), but can be useful as a quality control check to see if any surveys have numbers that are absolutely out of line in terms of the numbers of animals counted. Only certified or provisional data is included.

```{r echo=FALSE}

# Grab all the historical comp count data for the herd into a frame and show it
# Exclude raw data - provisional or certified only
Sql = paste("SELECT [Year], [Survey type], Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull, Adult, Caribou, Unknown, [Total groups observed] as n
-- ,CASE WHEN CertificationLevel = 'Certified' THEN '' ELSE '*' END As [Qual.]
FROM            Summary_Census
WHERE Herd = '",Herd,"' 
And [Survey type] = 'CC' 
And CertificationLevel <> 'Raw'
And Year <= ",Year,"
ORDER BY Year  DESC",sep="")
ComparisonDF = GetDataFrame(Sql)
kable(ComparisonDF)

# Plot the absolute numbers
# ggplot(data=ComparisonDF) +
#   geom_line(aes(x=Year,y = Calf, colour = "Calf")) + # I don't know why colour becomes the legend label, but it does
#   geom_line(aes(x=Year,y = Bull, colour = "Bull"))  +   
#   geom_line(aes(x=Year,y = Cow, colour = "Cow"))  +    
#   labs(title = "Sex and age composition, absolute numbers", color = "Categories\n") + 
#   theme_classic()

# Plot the calf/bull/cow ratios
CowsDF = ComparisonDF[,c('Cow')]
YMax = max(CowsDF, na.rm = TRUE)

plot(ComparisonDF$Year, ComparisonDF$Cow, type = "l", xlab = "Year", ylab = "Count",ylim = c(0,YMax), main="Absolute sex and age counts by category", sub=SurveyName )           
lines(ComparisonDF$Year, ComparisonDF$Bull, type = "l", col = "red")  
lines(ComparisonDF$Year, ComparisonDF$Calf, type = "l", col = "blue")  
legend("topright",legend = c("Cow","Bull","Calf"), col = c("black", "red","blue"),       lty = 1)

FigureCount = FigureCount + 1
```

Figure `r FigureCount`. Absolute cow counts over time.

# Appendix E: Available collars during the survey (data from Animal_Movement)

Appendix E contains a list of radiocollars that were available at the time of the survey according to the Animal_Movement database.

```{r echo=FALSE}

# Get a list of available collars from Animal_Movement (this stored procedure is executed through the WRST_Caribou database).
# The data will only be as current as the last time the data was pulled from Animal_Movement to WRST_Caribou by executing the spPullAMTables stored procedure
Sql = paste("exec dbo.GetListOfAvailableFrequenciesForAHerdAndDate '",Herd,"','",StartDate,"'",sep="")
QCDF = GetDataFrame(Sql)
# Subset out just the pertinent columns
AnimalsDF = QCDF[ ,c("AnimalId","Frequency","DeploymentDate","MortalityDate","RetrievalDate")]
```

Table `r TableCount`. Inventory of collars (`r nrow(QCDF)`) that were available during the survey time frame.

```{r echo=FALSE}

kable(AnimalsDF)

TableCount = TableCount + 1
```

```{r}
# Rename the output file to the Survey name
to = paste(SurveyName,' Summary.html')
file.copy(from='WRST-Caribou-Composition-Count-Survey-NRDS-Report.html',to=to)
```
