---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Setup
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(RODBC)
```



```{r echo=FALSE}

# Database connection string
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')

# Function that returns a data frame from an SQL query on the database
# SQL: A valid SQL query
GetDataFrame <- function(SQL) { # create a function with the name my_function
  DF = sqlQuery(Connection,SQL)
  return(DF)
}

# Use the below for sequentially numbering tables and figures
QCCheckCounter = 1
TableCounter = 1
```

# Data Release Report: Monitoring the Abundance, Composition and Survivorship of the Chisana and Mentasta Caribou Herds in Wrangell-St. Elias National Park and Preserve, Alaska

Scott D. Miller  
Information Technology Specialist/Data Manager  
National Park Service, Arctic Inventory and Monitoring Network

Generated by `r Sys.getenv("USERNAME")`
`r format(Sys.Date(), "%B %d, %Y")`

# Introduction

Abundance, composition and survivorship of the Chisana and Mentasta caribou herds have been monitored in Wrangell-St. Elias National Park and Preserve (WRST) for many decades ( [Putera and Miller, 2018](https://irma.nps.gov/DataStore/Reference/Profile/2256522), Table `r TableCounter`). Aerial surveys, aided by telemetry from GPS and VHF collars have provided a rich baseline of information about the herd. From 2017-2024 efforts were made to migrate caribou monitoring data from spreadsheets, files and paper data collection forms into a centralized, authoritative database and to apply modern data management structure and techniques to the dataset in order to preserve its quality, longevity and security into the future. The WRST caribou monitoring protocol requires data to undergo a disciplined set of steps from data collection, through database integration and quality control, to analysis and publication. At the end of the data processing life cycle a certified dataset suitable for analysis will be published to the NPS Data Store. This report is to accompany the certified dataset in order to provide context, document data defects and provide metadata that may be important for users of the dataset to know.

# Dataset Components

The WRST caribou certified dataset has two main components, 1) census surveys (herd population and composition surveys), and 2) radiotracking surveys. The objective of population surveys is to provide an estimate of the herd size. The size may be a minimum count, or it may have a sightability correction factor applied to it (Putera and Miller, 2018). Composition surveys provide sex and age ratios that may be used to infer the current and future structure of the herds. Radiotracking surveys provide information about longevity and survivorship of individual GPS collared animals, as well as opportunistic composition information. 

# Methods

Caribou monitoring data has been collected in a standard way, with some modifications over the years, since the 1960s. In 2018 the Central Alaska Inventory and Monitoring Network collaborated with WRST to develop a monitoring protocol (Putera and Miller, 2018), formalizing and describing all aspects of the caribou monitoring program at WRST. This protocol is available from the NPS Data Store at the following URL: [https://irma.nps.gov/DataStore/Reference/Profile/2256522](https://irma.nps.gov/DataStore/Reference/Profile/2256522). Digital reports, summaries and other products generated by the WRST caribou monitoring program are available at [https://www.nps.gov/im/cakn/caribou.htm](https://www.nps.gov/im/cakn/caribou.htm).

# Length Of Record By Herd And Survey Type

Table `r TableCounter`. WRST caribou monitoring dataset, length of record by herd and survey type. 
```{r echo=FALSE, label='Summary_LengthOfRecordByHerdAndSurveyType' }
Sql = "SELECT        Herd, [Survey type], Earliest, [Latest survey], [Years of record],[Survey days], Surveys as [Number of surveys]
FROM            Summary_LengthOfRecordByHerdAndSurveyType"
DF = GetDataFrame(Sql)
kable(DF)
QCCheckCounter = QCCheckCounter + 1
TableCounter = TableCounter + 1
```

# Current Dataset Certification Status

Dataset certification is the last step in the data management life cycle before analysis and provides assurance that the dataset has undergone quality control procedures, had defects remedied or documented, and is of analytical quality. Individual database records have a CertificatiorLevel attribute that is tagged according to the definition in Table `r TableCounter + 1`. Table `r TableCounter` shows a summary of the current dataset standing with respect to certification level by herd and survey type.

Table `r TableCounter`. Summary of certification level by herd and survey type.
```{r echo=FALSE, label='Summary_CensusSurveys_CountsByCertificationLevel' }
Sql = "SELECT * FROM [Summary_CensusSurveys_CountsByCertificationLevel]"
DF = GetDataFrame(Sql)
kable(DF)
QCCheckCounter = QCCheckCounter + 1
TableCounter = TableCounter + 1
```

Table `r TableCounter + 1`. Description of certification levels: `r TableCounter = TableCounter + 1`

| CertificationLevel | Description                                                                                                                                                                                                                                                                                            |
|--------------------|----------------------------------------------------|
| Raw                | Records imported directly from field surveys with no quality control or validation procedures performed.                                                                                                                                                                                               |
| Provisional        | Records that have undergone minimal quality control and/or have been examined by the project leader and are deemed suitable for internal reports but have not been validated against a published summary or report. Some records may never be fully certified due to the lack of validating materials. |
| Certified          | Records that have undergone extensive quality control procedures and have been validated against published reports, journal articles or data summary reports.                                                                                                                                          |



# Quality Control Checks

## Census Surveys (Population and Composition Surveys)

This portion of the quality control report is concerned with the population and composition surveys portion of the WRST caribou monitoring database. The checks in this section are related to the CaribouGroups data table and its hierarchically related GroupMembers data table.

### QC-`r QCCheckCounter`: Surveys having one or more records with no spatial information (Lat is NULL or Lat = 0).
```{r echo=FALSE, label='QC_CaribouGroups_SurveysHavingOneOrMoreRecordsWithBadSpatialCoordinates' }
Sql = 'SELECT      [SurveyName],  [Herd]
      ,[NumberOfRecordsMissingSpatialCoordinates]
  FROM [WRST_Caribou].[dbo].[QC_CaribouGroups_SurveysHavingOneOrMoreRecordsWithBadSpatialCoordinates]
ORDER By SurveyName'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  The quality control table below shows Surveys having one or more records with no spatial information (Lat is NULL or Lat = 0).  

Table QC-`r QCCheckCounter`. Surveys having one or more records with no spatial information (Lat is NULL or Lat = 0)  (`r nrow(DF)` records).
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.


### QC-`r QCCheckCounter`:  Surveys Having One Or More Records With No Counts
```{r echo=FALSE, label='QC_CaribouGroups_SurveysHavingOneOrMoreRecordsWithNoCounts' }
Sql = 'SELECT TOP (1000) [SurveyName]
      ,[NumberOfRecords]
  FROM [WRST_Caribou].[dbo].[QC_CaribouGroups_SurveysHavingOneOrMoreRecordsWithNoCounts]
ORDER By SurveyName'
DF = GetDataFrame(Sql)
```
Level: Advisory.  
**Description:** The records in the quality control table below have all animal count fields set to zero (no usable data). These are likely empty rows that were imported into the database accidentally.

Table QC-`r QCCheckCounter`. Surveys Having One Or More Records With No Counts  (`r nrow(DF)` records).
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.


### QC-`r QCCheckCounter`:  Surveys Having Frequencies Count Not Matching Available Frequencies In Animal_Movement
```{r echo=FALSE, label='QC_CaribouGroups_SurveysHavingFrequenciesCountNotMatchingAvailableFrequenciesInAnimal_Movement' }
Sql = 'SELECT [SurveyName] 
   --   ,[Survey type]
      ,[Herd]
      ,[Count of frequencies]
      ,[AvailableCollars] as [Available collars]
      --,[Year]
  FROM [WRST_Caribou].[dbo].[QC_CaribouGroups_SurveysHavingFrequenciesCountNotMatchingAvailableFrequenciesInAnimal_Movement]
ORDER By SurveyName'
DF = GetDataFrame(Sql)
```
Level: Advisory.  
**Description:** During aerial caribou surveys it is most often the goal to detect every GPS collar frequency available in the field. This goal is often not met due to animals straying out of the survey area, collar failures, terrain blocking receivers, collar inventory omissions in the Animal_Movement database, or other problems. The quality control table below shows surveys where there is a mismatch between the count of frequencies that were available for the herd and date of the survey according to the Animal_Movement database, and the number of frequencies that were detected in the field.   

Table QC-`r QCCheckCounter`. Surveys Having Frequencies Count Not Matching Available Frequencies In Animal_Movement  (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.





### QC-`r QCCheckCounter`: Surveys Having Frequencies Detected But [Marked] attribute Is NULL
```{r echo=FALSE, label='QC_CaribouGroups_SurveysWhereFrequenciesDetectedButMarkedIsNULL' }
Sql = 'SELECT * FROM QC_CaribouGroups_SurveysWhereFrequenciesDetectedButMarkedIsNULL ORDER BY SurveyName'
DF = GetDataFrame(Sql)
```
Level: Advisory.   
**Description:**  Sightability is an important component of population surveys where a correction factor must be developed statistically to account for animals that were available to be detected, but weren't. Each survey record has FrequenciesInGroup and Marked attributes. If a GPS collar frequency is detected then it is recorded in the FrequenciesInGroup column and the caribou group record's Marked attribute is set to 1. The table below shows surveys that may have records that should be examined for adherence to this rule. This is an advisory because other factors come into play. An animal may be detected, but outside the survey area in which case it is not important that it be [Marked] because it was unavailable for sightability purposes. Other mitigating factors may be at work. 

Table QC-`r QCCheckCounter`. Surveys Having Frequencies Detected But [Marked] attribute Is NULL  (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.




### QC-`r QCCheckCounter`: Surveys Having Frequencies Detected But No Group Member Records Exist
```{r echo=FALSE, label='QC_CaribouGroups_SurveysHavingFrequenciesDetectedButNoGroupMemberRecords' }
Sql = 'SELECT * FROM QC_CaribouGroups_SurveysHavingFrequenciesDetectedButNoGroupMemberRecords ORDER BY SurveyName'
DF = GetDataFrame(Sql)
```
**Level:** Critical  
**Description:**  Survey crews are to record all GPS collar frequencies they detect as comma separated values in the FrequenciesInGroup attribute of the caribou group record. These frequencies are later translated from frequencies to AnimalIDs as stored in the Animal_Movement GPS collar database. This quality control check is to ensure that the post-survey frequencies to AnimalIDs translation has occurred and every CaribouGroup row having non-NULL FrequenciesInGroup attribute has related records in the GroupMembers table with appropriate AnimalIDs.

Table QC-`r QCCheckCounter`. Surveys Having Frequencies Detected But No Group Member Records Exist  (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.



### QC-`r QCCheckCounter`: Unmatched Frequencies
```{r echo=FALSE, label='QC_CaribouGroups_UnmatchedFrequencies' }
Sql = 'SELECT * FROM QC_CaribouGroups_UnmatchedFrequencies'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  If a CaribouGroup's FrequenciesInGroup attribute is not NULL then there must be a record in the GroupMembers table with the frequency's value translated to an AnimalID. 

Table QC-`r QCCheckCounter`. Frequencies detected during an aerial census survey that have not been translated to a GroupMember record containing the frequency's translated AnimalID  (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.




## Radiotracking Surveys

This portion of the quality control report is concerned with the radiotracking surveys portion of the WRST caribou monitoring database. The checks in this section are related to the Radiotracking data table.


### QC-`r QCCheckCounter`: Animals observed after their mortality date.
```{r echo=FALSE, label='QC_Radiotracking_AnimalsMarkedDeadOnMultipleDates' }
Sql = 'SELECT * FROM QC_Radiotracking_AnimalsMarkedDeadOnMultipleDates'
DF = GetDataFrame(Sql)
```
**Level:** Critical.  
**Description:**  This view shows radiotracking records of animals observed after their mortality date as recorded in Animal_Movement.Animals table. The most common source of this error occurs when GPS collars drop to the ground after caribou mortality, or planned or accidental collar release and the collar is then re-detected on a subsequent survey. Occasionally the MortalityDate attribute for the caribou is erroneous, leading to data inconsistencies.

Table QC-`r QCCheckCounter`. Radiotracking records of animals observed after their mortality date as recorded in Animal_Movement.Animals  table  (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.

### QC-`r QCCheckCounter`: Incorrect AnimalID For DeploymentID
```{r echo=FALSE, label='QC_Radiotracking_IncorrectAnimalIDForDeploymentID' }
Sql = 'SELECT * FROM QC_Radiotracking_IncorrectAnimalIDForDeploymentID'
DF = GetDataFrame(Sql)
```
**Level:** Critical.  
**Description:**  Collar deployment details are stored in the NPS Animal_Movement database. Each collar deployment on an animal has a unique DeploymentID. This QC test shows records for which the DeploymentID may be incorrect.  

Table QC-`r QCCheckCounter`. Incorrect AnimalID For DeploymentID.  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.

### QC-`r QCCheckCounter`: Mode Recorded As 'M' (Mortality) But No Mortality Date In Animal_Movement
```{r echo=FALSE, label='QC_Radiotracking_ModeIsMButNoMortalityDateInAnimal_Movement' }
Sql = 'SELECT * FROM QC_Radiotracking_ModeIsMButNoMortalityDateInAnimal_Movement'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  GPS collars are programmed to transmit Mode = M when an animal dies in the field. The records in this QC test show Mode = M, but the animal has not been marked dead (MortalityDate Is Not NULL) in Animal_Movement database. One cause of this may be a dropped collar or the collar is found but it is ambiguous as to whether a mortality occurred.

Table QC-`r QCCheckCounter`. Mode Recorded As 'M' (Mortality) But No Mortality Date In Animal_Movement.  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.

### QC-`r QCCheckCounter`: Observed Versus Available Frequencies For A Survey Date
```{r echo=FALSE, label='QC_Radiotracking_ObservedVersusAvailableFrequenciesForASurveyDate' }
Sql = 'SELECT * FROM QC_Radiotracking_ObservedVersusAvailableFrequenciesForASurveyDate ORDER By SightingDate'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  During a radiotracking the goal is to find each collared animal that is available in the field. This QC checks shows radiotracking surveys where this goal was not met. This check is an advisory because it is common to miss collars for various reasons such as collar failure, terrain blocking signals or animals being out of range.  

Table QC-`r QCCheckCounter`. Observed Versus Available Frequencies For A Survey Date (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.


### QC-`r QCCheckCounter`:  Possible Adult Summing Errors
```{r echo=FALSE, label='QC_Radiotracking_PossibleAdultSummingErrors' }
Sql = 'SELECT * FROM QC_Radiotracking_PossibleAdultSummingErrors'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  Each CaribouGroup record's Adult attribute should reflect the total of bulls and cows. The records shown below violate this rule.   

Table QC-`r QCCheckCounter`. Possible Adult Summing Errors (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.


### QC-`r QCCheckCounter`: Possible Total Caribou Summing Errors
```{r echo=FALSE, label='QC_Radiotracking_PossibleTotalCaribouSummingErrors' }
Sql = 'SELECT * FROM QC_Radiotracking_PossibleTotalCaribouSummingErrors'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  Each CaribouGroup record's Caribou attribute should reflect the total of bulls, calves and cows. The records shown below violate this rule.  

Table QC-`r QCCheckCounter`. Possible Total Caribou Summing Errors (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.


### QC-`r QCCheckCounter`: Possibly Incorrect AnimalID
```{r echo=FALSE, label='QC_Radiotracking_PossiblyIncorrectAnimalID' }
Sql = 'SELECT * FROM QC_Radiotracking_PossiblyIncorrectAnimalID'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  Possibly Incorrect AnimalID. The records shown below have a DeploymentID that does not match the AnimalID in the Animal_Movement database.

Table QC-`r QCCheckCounter`. Possibly Incorrect AnimalID (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.


### QC-`r QCCheckCounter`: This query returns radiotracking records where a caribou was observed on a date following its recorded MortalityDate in the Animal_Movement.Animal table.
```{r echo=FALSE, label='QC_Radiotracking_PostMortalitySightings' }
Sql = 'SELECT * FROM QC_Radiotracking_PostMortalitySightings'
DF = GetDataFrame(Sql)
```
**Level:** Critical.  
**Description:**  This query returns radiotracking records where a caribou was observed on a date following its recorded MortalityDate in the Animal_Movement.Animal table..  

Table QC-`r QCCheckCounter`. This query returns radiotracking records where a caribou was observed on a date following its recorded MortalityDate in the Animal_Movement.Animal table (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.



### QC-`r QCCheckCounter`: Unreconciled Mortality Dates With Animal_Movement

```{r echo=FALSE, label='QC_Radiotracking_UnreconciledMortalityDatesWithAnimal_Movement' }
Sql = 'SELECT * FROM QC_Radiotracking_UnreconciledMortalityDatesWithAnimal_Movement'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  Mortalities are recorded in the field during radiotracking surveys. This view should help determine if the field recorded mortalities reconcile with mortalities in Animal_Movement. Keep in mind the mortality date in Animal_Movement may be more accurate than what is in the WRST_Caribou.Radiotracking table because the mortality date may have been determined to the hour from a GPS collar while an older mort date from a VHF collar may not be known to an accuracy of months between surveys. Use judgement in assigning mortality dates in Animal_Movement. This tool is probably most useful for ensuring that dead animals are recorded dead in both databases.   

Table QC-`r QCCheckCounter`. Unreconciled Mortality Dates With Animal_Movement  (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.

### QC-`r QCCheckCounter`: Deat Caribou Resighted.
```{r echo=FALSE, label='QC_Surveys_DeadCaribouResighted' }
Sql = 'SELECT * FROM QC_Surveys_DeadCaribouResighted'
DF = GetDataFrame(Sql)
```
**Level:** Advisory.  
**Description:**  This test reveals any caribou that were observed on a survey that were prior recorded as dead. This is most likely an observer detecting a collar from a dead caribou that was left in the field, but may be an incorrectly recorded mortality date in the Animal_Movement.Animals table..  

Table QC-`r QCCheckCounter`. This test reveals any caribou that were observed on a survey that were prior recorded as dead. This is most likely an observer detecting a collar from a dead caribou that was left in the field, but may be an incorrectly recorded mortality date in the Animal_Movement.Animals table (`r nrow(DF)` records).  
```{r echo=FALSE }
if (nrow(DF) == 0) {
    Result <- c('PASSED: No records returned')
    DF <- data.frame(Result)
}
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```
Number of records: `r nrow(DF)`.  
**Query:** _`r Sql`_.

# Appendix A: Database Metadata

## Data Tables Descriptions

Table A-1: Database tables descriptions.  
```{r echo=FALSE, label='DatabaseTablesDescriptions' }
Sql = "select [Table],TableDescription as [Description] from DatabaseTableDescriptions where type='U'
and [Table] in ('CaribouGroups','GroupMembers','Radiotracking')"
DF = GetDataFrame(Sql)
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```

## Columns Descriptions

Table A-2: Columns Descriptions, Data Types and Units of Measure.  
```{r echo=FALSE, label='DatabaseColumnsDescriptions' }
Sql = "SELECT        [Table], [Column], ColumnDescription As Description
FROM            DatabaseColumnsDescriptions
WHERE        ([Table] IN ('CaribouGroups', 'GroupMembers', 'Radiotracking'))
ORDER BY [Table]"
DF = GetDataFrame(Sql)
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```

# Appendix B: Work Log

Table A-2: Columns Descriptions, Data Types and Units of Measure.  
```{r echo=FALSE, label='WorkLog' }
Sql = "SELECT        LogDate, Author, LogEntry
FROM            WorkLog
ORDER BY LogDate DESC"
DF = GetDataFrame(Sql)
kable(DF)
QCCheckCounter = QCCheckCounter + 1
```

# References

Putera JA and Miller SD. 2018. Protocol for monitoring caribou populations in Wrangell-St. Elias National Park & Preserve, Central Alaska Network: Narrative – version 1.0. Natural Resource Report. NPS/CAKN/NRR—2018/1750. National Park Service. Fort Collins, Colorado









