```{r echo=FALSE}
# Set report parameters
Year = 2024 # Year 
Park = 'WRST'
```

---
title: "`r Year ` Annual Report: Monitoring the Abundance and Composition of the Chisana and Mentasta Caribou Herds in Wrangell - St. Elias National Park and Preserve, Alaska"
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, include=FALSE}

# Setup
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(RODBC)
library(sf)
library(ggplot2)
library(SightabilityModel)
library(tidyverse)

# WRST Shortcut
WRST = "Wrangell - St. Elias National Park and Preserve"

# Database connection string
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')

# Counters to increment the table and figure numbers as the report progresses.
TableCount = 1
FigureCount = 1

```

Kyle Cutting\
Wildlife Biologist\
National Park Service, `r WRST`\
8 Richardson Hwy Mile 106\
Copper Center, AK 99573

Scott D. Miller\
Information Technology Specialist/Data Manager\
National Park Service, Arctic Inventory and Monitoring Network\
240 W. 5th Ave.\
Anchorage, AK 99501

# Abstract

# Introduction

## Chisana Surveys History

### 1987-2002
Chisana 87-2002 pop est used late june minimum counts, and fall sex and age composition, recruitment and mortality data using adfg model

### 2003-2006
Captive rearing 2003-6. All animals collared and released. (208 animals). USGS implemented fall censos technique correcting for undetected caribou using radiocollared animals.

# Objectives

For each herd monitor
-   Sex ratios  
-   Calf recruitment  
-   Population size  

## Mentasta Surveys History


# Methods

# Results

## Chisana

### Herd Distribution Map

Caribou were distributed as shown in Figure `r FigureCount`.

```{r echo=FALSE}

# Get the spatial data for comp and population surveys
Sql = paste("SELECT        SurveyName, Year, Herd, SurveyType, SearchArea, GroupNumber, SightingDate, Lat, Lon, [In], Seen, Marked
FROM            CaribouGroups
WHERE (Year = ",Year,") And Herd='Chisana' And Lat > 0",sep="")
data = sqlQuery(Connection,Sql)

# Convert the data into a spatial data frame
spdata <- sf::st_as_sf(data, coords = c('Lon','Lat'))

# Set the coordinate system to GCS Lat\Lon WGS1984
st_crs(spdata) <- st_crs(4326)

# Park boundaries (geojson) are at https://irma.nps.gov/DataStore/Reference/Profile/2303652
# Create a temporary file on user's hard drive
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the spatial data file to the temporary file location created above
download.file("https://irma.nps.gov/DataStore/DownloadFile/702822",tmp_geojson)

# Read the temporary spatial data file into an sf object
AKParks = read_sf(tmp_geojson)

# Show a map of where caribou groups were found during the survey. Generate a density heat map also.
#Plot the points over the shapefile (shapefile first, then points)
ggplot() + 
  
  # Park boundary
  geom_sf(data = AKParks,fill = "white", color = "black") + 
  
  # Caribou kernel density
  geom_density_2d(data = data, aes(x = Lon, y = Lat), colour = "red", alpha = 0.9) +
  
  # Caribou observations
  geom_sf(data = spdata, aes(color=factor(Marked)), size= 1, show.legend = T) +
  scale_color_manual(values = c("black","darkgray"))  +
  guides(color=guide_legend(title="Marked (1=True,0=False)")) +
  
  # Adjust colors
  scale_fill_brewer() + # This seems to make the colors less wacky, not sure why
  guides(fill = guide_legend(title = "Density of observations")) + # This gives a legend to the density
  
  # Limit the view port to just WRST
  xlim(-146.5,-139) +
  ylim(59.8,62.65) +
  #ggtitle(paste("Caribou observations distribution and density",sep="") ,subtitle = CompDF$`Start date`) +
  
  # Label axes
  xlab('Longitude') + ylab('Latitude') +
  
  ggtitle('Distribution of Chisana herd during survey') +
  
  # Clean look
  theme_classic()

```

### Results

Table X. Chisana caribou survey observations summary.

```{r echo=FALSE}

# Results of the Chisana Comp. Survey - Raw counts
Sql = paste("SELECT 
Cow
, Calf
, [Small bull]
, [Medium bull]
, [Large bull]
, Bull, Adult
, Caribou
, IsNull(Unknown,0) as Unknown
, [Minimum group size]
, [Maximum group size]
, [Total groups observed]
, Marked
, Seen
, [Total collared animals observed]
, [Number of frequencies that were available on survey date (Animal_Movement)]
FROM            Summary_Census
WHERE        (Herd = 'Chisana') AND (Year = ",Year,")",sep="")
data = sqlQuery(Connection,Sql)

# Transpose and display the records
if (nrow(data) > 0){
  # If we got rows show them
  colnames(data_t) = "Result"
  kable(t(data))
}else{
  # Otherwise indicate no rows returned
  kable("No records")
}


```

Table X. Summary statistics

```{r echo=FALSE}
# Results of the Chisana Comp. Survey - Stats
Sql = paste("SELECT 
[Calves/100 cows]
, [Bulls/100 cows]
, [Pct. calf]
, [Pct. cow]
, [Pct. bull (composition survey)] as [Pct. bull]
, [Pct. small bull]
, [Pct. medium bull]
, [Pct. large bull]
, [Pct. small bull (of total bulls)]
, [Pct. medium bull (of total bulls)]
, [Pct. large bull (of total bulls)]
, GroupSize_Mean

--, [Type of survey], Timing,[Survey name]
FROM            Summary_Census
WHERE        (Herd = 'Chisana') AND (Year = ",Year,")",sep="")

# Transpose and display the records
if (nrow(data) > 0){
  # If we got rows show them
  colnames(data_t) = "Result"
  kable(t(data))
}else{
  # Otherwise indicate no rows returned
  kable("No records")
}


```

### Population Estimate

Note: 2024-06-27 SDMiller: I'm still working out how to run this analysis by trying to reproduce the 2010 Chisana Report put out by Yukon (https://irma.nps.gov/DataStore/DownloadFile/566074)[https://irma.nps.gov/DataStore/DownloadFile/566074]
Example code at J:\Monitoring\Caribou\WRST\Historical Files from Copper Center\Electronic\Chisana\SightabilityModel

```{r echo=FALSE}
# Observational dataset
# Note: For some reason Total and Grp_Size are both needed by SightabilityModel R package, even though they are duplicative
Sql = paste("SELECT  1 as stratum,[Total] as total, 1 as subunit,[Grp_Size] FROM [WRST_Caribou].[dbo].[Sightability_odat] WHERE Herd='Chisana' and Year = ",Year," And [In] = 1",sep="")
odat = sqlQuery(Connection,Sql)

# Sightability dataset
Sql = paste("SELECT   ID, Observed, Grp_Size FROM  Sightability_sdat WHERE (Herd = 'Chisana') AND (Year = ",Year,")",sep="")
sdat = sqlQuery(Connection,Sql)

# Sampling info dataset (essentially all 1s, generate on the fly)
stratum <- 1
Nh = 1
nh = 1
sampinfo <- data.frame(stratum,Nh,nh)

# If there are observation data records then fit a sightability model
# Transpose and display the records
if (nrow(odat) > 0){
  #Code to fit a sightability model using group size as a covariate
  fit.group.size <- Sight.Est(form=Observed ~ Grp_Size,sdat=sdat,odat=odat,sampinfo=sampinfo, method="Wong", logCI=T, alpha=0.1, Vm.boot=T, nboot=1000)
  fit.group.size #Provides the results from the model fitting
  
  #Code to fit a sightability model having a constant sightability (assumes 100% sightability?)
  fit.constant<-Sight.Est(form=Observed ~ 1, sdat=sdat, odat=odat, sampinfo=sampinfo, method="Wong", logCI=T, alpha=0.1, Vm.boot=T, nboot=1000)
  fit.constant  #Provides the results
  
  #The AICc for the constant model is lower than for the Group Size model so I chose the contant model as most supported.
  # Sight.Est doesn't calculate AICc (the small sample correction for AIC), I did that myself.
  
  #To get the actual sightability correction factor use the following codes
  #This fits the sightability model
  fit.glm <- glm(Observed~1, data=sdat, family=binomial()) 
  
  #This gives you the sightability correction factor (theta)
  theta.SS <- 1+exp(-coef(fit.glm)-summary(fit.glm)$cov.unscaled/2) 
  theta.SS # Provides the actual number

}else{
  # Otherwise indicate no rows returned
  kable("Observation dataset is empty")
}

```


### Sex and Age Ratios

One of the purposes of the management plan is to allow a small harvest provided the herd remains stable and meets a minimum ratio of ≥ 35 bulls per 100 cows each year, and an observed 3-year average of ≥ 15 calves per 100 cows

```{r echo=FALSE}
Sql = "SELECT        Herd,Year, [Survey name], [Survey type], [Type of survey], Timing, Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull, BullsWereCategorized, Adult, Caribou, Unknown, [Calves/100 cows], [Bulls/100 cows]
FROM            Summary_Census
WHERE Herd='Chisana' And [Survey type]='CC'
ORDER BY Herd,Year, [Survey name]"

data = sqlQuery(Connection,Sql)

ggplot(data) +
  geom_line(aes(x=Year,y=`Bulls/100 cows`,linetype = 'Bulls/100 cows')) +
  geom_line(aes(x=Year,y=`Calves/100 cows`,linetype = 'Calves/100 cows')) +
  geom_hline(yintercept=35, linetype='solid')+
  geom_hline(yintercept=15, linetype='solid')+

  guides(linetype=guide_legend(title="")) +

  # Make X axis labels vertical
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  
  theme_minimal() +  
  
  # Y axis label
  ylab("Ratio") +
  
  # X axis label
  xlab("Year") 
```

Figure X. Sex and age ratios over time for the Chisana caribou herd. The top solid horizontal line represents the goal bulls/100 cows ratio, and the bottom solid horizontal line represents the goal calves/100 cows ratio.

Table `r TableCount`. Census surveys accomplished in `r Year`.

```{r echo=FALSE}
# List out the surveys accomplished
Sql = paste("SELECT [Survey name],Herd, [Start date], [End date], [Type of survey], [Total groups observed], [Survey days], [Search areas]
FROM            Summary_Census
WHERE        (Year = ",Year,") And Park = '",Park,"'
ORDER BY [Start date] ",sep="")
SurveysDoneDF = sqlQuery(Connection,Sql)

kable(SurveysDoneDF[,c("Start date","End date","Herd","Type of survey","Total groups observed","Survey days","Search areas")])
TableCount = TableCount + 1

```

Table `r TableCount`. Radiotracking surveys accomplished in `r Year`.

```{r echo=FALSE}
# List out the surveys accomplished
Sql = paste("SELECT TOP (1000) [Herd]
      ,[Survey date]
      ,[Animals tracked]
      ,[Total frequencies available on this date]
      ,[Mortalities]
      ,[Active]
      ,[Mode unknown]
      ,[Pct mortality]
  FROM [WRST_Caribou].[dbo].[Summary_Radiotracking] 
  WHERE YEAR([Survey date])=",Year," ORDER BY [Survey date]",sep="")
SurveysDoneDF = sqlQuery(Connection,Sql)
kable(SurveysDoneDF)
TableCount = TableCount + 1
```

Table `r TableCount`. Flight details, `r Year`.

```{r echo=FALSE}
# Get the survey days and observers into a frame
Sql = paste("SELECT  DISTINCT Herd,
CASE WHEN SurveyType = 'PE' THEN 'Population' WHEN SurveyType='CC' THEN 'Composition' END As [Survey type]
,   SearchArea, CONVERT(Date, SightingDate) AS [Date], Observer1 AS Observer, Pilot
FROM            CaribouGroups
WHERE         (Year = ",Year,") And Park = '",Park,"'
ORDER BY Herd,[Survey type],CONVERT(Date, SightingDate),SearchArea, Observer, Pilot"
,sep="")
ObsPilotsDF = sqlQuery(Connection,Sql)
kable(ObsPilotsDF)
TableCount = TableCount + 1
```

# Discussion
