```{r echo=FALSE}
# SET USER DEFINED PARAMETERS HERE
Year = 2024 # Year 
Herd = 'Mentasta'
SurveyType = 'PE'

```

```{r echo=FALSE}
# Park shortcuts
Park='WRST'
WRST = "Wrangell - St. Elias National Park and Preserve"

# Convert the survey type acronym into a longer form for use in the text
SurveyTypeLong=""
if(SurveyType=='PE'){
  SurveyTypeLong='Population Estimate'
} else if(SurveyType=='CC'){
    SurveyTypeLong='Population Estimate'
}
Title = paste("Survey Summary: ",Year," ",SurveyTypeLong," of the ",Herd," Caribou Herd in ",WRST,", Alaska",sep="")
```

---
title: "`r Title`" 
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
---

```{r setup, echo=FALSE, include=FALSE}

# Setup
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(RODBC)
library(sf)
library(ggplot2)
library(SightabilityModel)
library(tidyverse)



# Database connection strings
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')
AM_Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpakrovmais;Database=Animal_Movement;Trusted_connection=yes')

# Counters to increment the table and figure numbers as the report progresses.
TableCount = 1
FigureCount = 1

```


Scott D. Miller\
Information Technology Specialist/Data Manager\
National Park Service, Arctic Inventory and Monitoring Network\
240 W. 5th Ave.\
Anchorage, AK 99501

# Results

## Herd Distribution

Caribou were distributed as shown in Figure `r FigureCount`.

```{r echo=FALSE,warning=FALSE}

# Get the spatial data for comp and population surveys
Sql = paste("SELECT SurveyName, Year, Herd, SurveyType, SearchArea, GroupNumber, SightingDate, Lat, Lon, [In], Seen
,CASE When Marked = 1 Then 'True' WHEN Marked = 0 Then 'False' END as Marked
--, Marked
FROM Dataset_Census
WHERE Year = ",Year," And Herd='",Herd,"' And Lat > 0",sep="")
data = sqlQuery(Connection,Sql)

# Convert the data into a spatial data frame
spdata <- sf::st_as_sf(data, coords = c('Lon','Lat'))

# Set the coordinate system to GCS Lat\Lon WGS1984
st_crs(spdata) <- st_crs(4326)

# Park boundaries (geojson) are at https://irma.nps.gov/DataStore/Reference/Profile/2303652
# Create a temporary file on user's hard drive
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the spatial data file to the temporary file location created above
download.file("https://irma.nps.gov/DataStore/DownloadFile/702822",tmp_geojson)

# Read the temporary spatial data file into an sf object
AKParks = read_sf(tmp_geojson)

# Show a map of where caribou groups were found during the survey. Generate a density heat map also.
#Plot the points over the shapefile (shapefile first, then points)
ggplot() + 
  
  # Park boundary
  geom_sf(data = AKParks,fill = "white", color = "black") + 
  
  # Caribou kernel density
  geom_density_2d(data = data, aes(x = Lon, y = Lat), colour = "red", alpha = 0.9) +
  
  # Caribou observations
  geom_sf(data = spdata, aes(color=factor(Marked)), size= 1, show.legend = T) +
  scale_color_manual(values = c("darkgray","black"))  +
  guides(color=guide_legend(title="Marked")) +
  
  # Adjust colors
  scale_fill_brewer() + # This seems to make the colors less wacky, not sure why
  guides(fill = guide_legend(title = "Density of observations")) + # This gives a legend to the density
  
  # Limit the view port to just WRST
  xlim(-146.5,-139) +
  ylim(59.8,62.65) +
  #ggtitle(paste("Caribou observations distribution and density",sep="") ,subtitle = CompDF$`Start date`) +
  
  # Label axes
  xlab('Longitude') + ylab('Latitude') +
  
  ggtitle(paste("Distribution of ",Herd," herd during survey",sep="")) +
  
  # Clean look
  theme_classic()

```

```{r echo=FALSE,warning=FALSE}
# Survey summary
Sql = paste("SELECT 
[Cow]
,[Calf]
,[Small bull]
,[Medium bull]
,[Large bull]
,[Bull]
,[Adult]
,[Caribou]
,[Unknown]
,[Calf, male]
,[Calf, female]
,[Calf, unclassified]
,[Yearling]
,[GroupSize_Mean]
,[Minimum group size]
,[Maximum group size]
,[Pct. calf]
,[Pct. cow]
,[Pct. small bull]
,[Pct. medium bull]
,[Pct. large bull]
,[Pct. small bull (of total bulls)]
,[Pct. medium bull (of total bulls)]
,[Pct. large bull (of total bulls)]
,[Pct. bull (composition survey)]
,[Pct. bull (population survey)]
,[Calves/100 cows]
,[Bulls/100 cows]
,[Total groups observed]
,[Marked]
,[Seen]
,[Marked and Seen]
,[Marked and Not Seen]
,[Pct. marked groups observed]
,[Groups in the survey area]
,[Groups out of the survey area]
,[Marked groups in survey area]
,[Marked groups observed in survey area]
,[Observed groups in survey area]
,[Pct. marked groups observed in survey area]
,[Total collared animals observed]
,[Number of frequencies that were available on survey date (Animal_Movement)]
,[Collars not heard or not searched]
,[Start date]
,[End date]
,[Survey days]
,[Search areas]
,[Park]
,[Year]
--,[Survey name]
--,[Herd]
--,[Survey type]
--,[Type of survey]
--,Timing
--,[BullsWereCategorized]
FROM [WRST_Caribou].[dbo].[Summary_Census]
WHERE        (Herd = '",Herd,"') AND (Year = ",Year,") And [Survey type]='",SurveyType,"'",sep="")
Summary = sqlQuery(Connection,Sql)

```

## Counts

```{r echo=FALSE,warning=FALSE}

# Select out the counts only
Counts = Summary %>% select(Cow
,`Calf`
,`Small bull`
,`Medium bull`
,`Large bull`
,`Bull`
,`Adult`
,`Caribou`
,`Unknown`
,`Yearling`
,`Minimum group size`
,`Maximum group size`)
CountsDF = as.data.frame(Counts)
rownames(CountsDF) = NULL
knitr::kable(t(CountsDF))
```




### Results


```{r echo=FALSE}

# Summary of the survey
if (nrow(Summary) > 0){
  # If we got rows show them
  Summary_t = t(Summary)
  colnames(Summary_t) = "Result"
  kable(Summary_t)
}else{
  # Otherwise indicate no rows returned
  kable("No records")
}


```

Table X. Summary statistics

```{r echo=FALSE}
# Results of the  Comp. Survey - Stats
Sql = paste("SELECT 
[Calves/100 cows]
, [Bulls/100 cows]
, [Pct. calf]
, [Pct. cow]
, [Pct. bull (composition survey)] as [Pct. bull]
, [Pct. small bull]
, [Pct. medium bull]
, [Pct. large bull]
, [Pct. small bull (of total bulls)]
, [Pct. medium bull (of total bulls)]
, [Pct. large bull (of total bulls)]
, GroupSize_Mean

--, [Type of survey], Timing,[Survey name]
FROM            Summary_Census
WHERE        (Herd = '",Herd,"') AND (Year = ",Year,")",sep="")

# Transpose and display the records
if (nrow(data) > 0){
  # If we got rows show them
  colnames(data_t) = "Result"
  kable(t(data))
}else{
  # Otherwise indicate no rows returned
  kable("No records")
}


```

### Population Estimate

Note: 2024-06-27 SDMiller: I'm still working out how to run this analysis by trying to reproduce the 2010 Chisana Report put out by Yukon (https://irma.nps.gov/DataStore/DownloadFile/566074)[https://irma.nps.gov/DataStore/DownloadFile/566074]
Example code at J:\Monitoring\Caribou\WRST\Historical Files from Copper Center\Electronic\Chisana\SightabilityModel

```{r echo=FALSE}
# Observational dataset
# Note: For some reason Total and Grp_Size are both needed by SightabilityModel R package, even though they are duplicative
Sql = paste("SELECT  1 as stratum,[Total] as total, 1 as subunit,[Grp_Size] FROM [WRST_Caribou].[dbo].[Sightability_odat] WHERE Herd='",Herd,"' and Year = ",Year," And [In] = 1",sep="")
odat = sqlQuery(Connection,Sql)

# Sightability dataset
Sql = paste("SELECT   ID, Observed, Grp_Size FROM  Sightability_sdat WHERE (Herd = '",Herd,"') AND (Year = ",Year,")",sep="")
sdat = sqlQuery(Connection,Sql)

# Sampling info dataset (essentially all 1s, generate on the fly)
stratum <- 1
Nh = 1
nh = 1
sampinfo <- data.frame(stratum,Nh,nh)

# If there are observation data records then fit a sightability model
# Transpose and display the records
if (nrow(odat) > 0){
  #Code to fit a sightability model using group size as a covariate
  fit.group.size <- Sight.Est(form=Observed ~ Grp_Size,sdat=sdat,odat=odat,sampinfo=sampinfo, method="Wong", logCI=T, alpha=0.1, Vm.boot=T, nboot=1000)
  fit.group.size #Provides the results from the model fitting
  
  #Code to fit a sightability model having a constant sightability (assumes 100% sightability?)
  fit.constant<-Sight.Est(form=Observed ~ 1, sdat=sdat, odat=odat, sampinfo=sampinfo, method="Wong", logCI=T, alpha=0.1, Vm.boot=T, nboot=1000)
  fit.constant  #Provides the results
  
  #The AICc for the constant model is lower than for the Group Size model so I chose the contant model as most supported.
  # Sight.Est doesn't calculate AICc (the small sample correction for AIC), I did that myself.
  
  #To get the actual sightability correction factor use the following codes
  #This fits the sightability model
  fit.glm <- glm(Observed~1, data=sdat, family=binomial()) 
  
  #This gives you the sightability correction factor (theta)
  theta.SS <- 1+exp(-coef(fit.glm)-summary(fit.glm)$cov.unscaled/2) 
  theta.SS # Provides the actual number

}else{
  # Otherwise indicate no rows returned
  kable("Observation dataset is empty")
}

```


### Sex and Age Ratios

One of the purposes of the management plan is to allow a small harvest provided the herd remains stable and meets a minimum ratio of ≥ 35 bulls per 100 cows each year, and an observed 3-year average of ≥ 15 calves per 100 cows

```{r echo=FALSE}
Sql = "SELECT        Herd,Year, [Survey name], [Survey type], [Type of survey], Timing, Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull, BullsWereCategorized, Adult, Caribou, Unknown, [Calves/100 cows], [Bulls/100 cows]
FROM            Summary_Census
WHERE Herd='",Herd,"' And [Survey type]='CC'
ORDER BY Herd,Year, [Survey name]"

data = sqlQuery(Connection,Sql)

ggplot(data) +
  geom_line(aes(x=Year,y=`Bulls/100 cows`,linetype = 'Bulls/100 cows')) +
  geom_line(aes(x=Year,y=`Calves/100 cows`,linetype = 'Calves/100 cows')) +
  geom_hline(yintercept=35, linetype='solid')+
  geom_hline(yintercept=15, linetype='solid')+

  guides(linetype=guide_legend(title="")) +

  # Make X axis labels vertical
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  
  theme_minimal() +  
  
  # Y axis label
  ylab("Ratio") +
  
  # X axis label
  xlab("Year") 
```

Figure X. Sex and age ratios over time for the `r Herd` caribou herd. The top solid horizontal line represents the goal bulls/100 cows ratio, and the bottom solid horizontal line represents the goal calves/100 cows ratio.



Table `r TableCount`. Radiotracking surveys accomplished in `r Year`.

```{r echo=FALSE}
# List out the surveys accomplished
Sql = paste("SELECT TOP (1000) [Herd]
      ,[Survey date]
      ,[Animals tracked]
      ,[Total frequencies available on this date]
      ,[Mortalities]
      ,[Active]
      ,[Mode unknown]
      ,[Pct mortality]
  FROM [WRST_Caribou].[dbo].[Summary_Radiotracking] 
  WHERE YEAR([Survey date])=",Year," ORDER BY [Survey date]",sep="")
SurveysDoneDF = sqlQuery(Connection,Sql)
kable(SurveysDoneDF)
TableCount = TableCount + 1
```

Table `r TableCount`. Flight details, `r Year`.

```{r echo=FALSE}
# Get the survey days and observers into a frame
Sql = paste("SELECT  DISTINCT Herd,
CASE WHEN SurveyType = 'PE' THEN 'Population' WHEN SurveyType='CC' THEN 'Composition' END As [Survey type]
,   SearchArea, CONVERT(Date, SightingDate) AS [Date], Observer1 AS Observer, Pilot
FROM            CaribouGroups
WHERE         (Year = ",Year,") And Park = '",Park,"'
ORDER BY Herd,[Survey type],CONVERT(Date, SightingDate),SearchArea, Observer, Pilot"
,sep="")
ObsPilotsDF = sqlQuery(Connection,Sql)
kable(ObsPilotsDF)
TableCount = TableCount + 1
```

# Discussion


```{r, echo=FALSE}
# Rename the resulting html file with the year, herd and survey type
NewFilename = paste(Year," ",Herd," Caribou ",SurveyTypeLong," Survey Summary.html",sep='')
file.rename('WRST Caribou Census Summary.html', NewFilename)
```

