```{r echo=FALSE}

# Set report parameters
Year = 2021 # Year of the comp count survey
Park = 'WRST'
```

---
title: "Annual Report: Monitoring the Abundance and Composition of the Chisana and Mentasta Caribou Herds in Wrangell - St. Elias National Park and Preserve, Alaska"
author: "Scott D. Miller"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: yes
---

```{r setup, include=FALSE}

# Setup
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(RODBC)
library(sf)
library(ggplot2)

# WRST Shortcut
WRST = "Wrangell - St. Elias National Park and Preserve"

# Database connection string
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')

# Function that returns a data frame from an SQL query on the database
# SQL: A valid SQL query
GetDataFrame <- function(SQL) { # create a function with the name my_function
  DF = sqlQuery(Connection,SQL)
  return(DF)
}


# Counters to increment the table and figure numbers as the report progresses.
TableCount = 1
FigureCount = 1

```

Scott D. Miller\
Information Technology Specialist/Data Manager\
National Park Service, Arctic Inventory and Monitoring Network\
240 W. 5th Ave.\
Anchorage, AK 99501

Kyle Cutting\
Wildlife Biologist\
National Park Service, `r WRST`\
8 Richardson Hwy Mile 106\
Copper Center, AK 99573




# Abstract

# Introduction

# Methods

# Results



Table `r TableCount`. Census surveys accomplished in `r Year`.

```{r echo=FALSE}
# List out the surveys accomplished
SurveysDoneDFSQL = paste("SELECT [Survey name],Herd, [Start date], [End date], [Type of survey], [Total groups observed], [Survey days], [Search areas]
FROM            Summary_Census
WHERE        (Year = ",Year,") And Park = '",Park,"'
ORDER BY [Start date] ",sep="")
SurveysDoneDF = GetDataFrame(SurveysDoneDFSQL)

kable(SurveysDoneDF[,c("Start date","End date","Herd","Type of survey","Total groups observed","Survey days","Search areas")])
TableCount = TableCount + 1






```


Table `r TableCount`. Radiotracking surveys accomplished in `r Year`.

```{r echo=FALSE}
# List out the surveys accomplished
SurveysDoneDFSQL = paste("SELECT TOP (1000) [Herd]
      ,[Survey date]
      ,[Animals tracked]
      ,[Total frequencies available on this date]
      ,[Mortalities]
      ,[Active]
      ,[Mode unknown]
      ,[Pct mortality]
  FROM [WRST_Caribou].[dbo].[Summary_Radiotracking] 
  WHERE YEAR([Survey date])=",Year," ORDER BY [Survey date]",sep="")
SurveysDoneDF = GetDataFrame(SurveysDoneDFSQL)
kable(SurveysDoneDF)
TableCount = TableCount + 1
```







## Who did it
Table `r TableCount`. Flight details, `r Year`.

```{r echo=FALSE}
# Get the survey days and observers into a frame
Sql = paste("SELECT  DISTINCT Herd,
CASE WHEN SurveyType = 'PE' THEN 'Population' WHEN SurveyType='CC' THEN 'Composition' END As [Survey type]
,   SearchArea, CONVERT(Date, SightingDate) AS [Date], Observer1 AS Observer, Pilot
FROM            CaribouGroups
WHERE         (Year = ",Year,") And Park = '",Park,"'
ORDER BY Herd,[Survey type],CONVERT(Date, SightingDate),SearchArea, Observer, Pilot"
,sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)
TableCount = TableCount + 1
```

# Discussion
