
```{r echo=FALSE}

# Set report parameters
Year = 2022 # Year of the comp count survey
Herd = 'Chisana' # Herd
SurveyName = '2022 Fall Chisana Caribou Composition Survey' # Change this to the survey name in the WRST Caribou database.
```

---
title: Results from the `r Year` Composition Count Survey of the `r Herd` Caribou Herd in Wrangell - St. Elias National Park and Preserve, Alaska
author: "Scott D. Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}

# Setup
options(knitr.kable.NA = '')
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(RODBC)
library(sf)
library(ggplot2)

# I hate writing this out - create a variable for it
WRST = "Wrangell - St. Elias National Park and Preserve"

# Database connection string
Connection =  odbcDriverConnect('Driver={SQL Server};Server=inpyugamsvm01\\nuna;Database=WRST_Caribou;Trusted_connection=yes')

# Function that returns a data frame from an SQL query on the database
# SQL: A valid SQL query
GetDataFrame <- function(SQL) { # create a function with the name my_function
  DF = sqlQuery(Connection,SQL)
  return(DF)
}

# Function to convert a data frame of Lat/Lon pairs to a spatial data frame
# DataFrame: A data frame containing a latitude field and a longitude field
# LatitudeFieldName: The name of the field containing latitude data
# LongitudeFieldName: The name of the field containing longitude data
GetPointSpatialDataFrameFromDataFrame = function(DataFrame,LatitudeFieldName,LongitudeFieldName){
  
  # Convert the data frame into a spatial data frame
  SDF <- sf::st_as_sf(DataFrame, coords = c(LongitudeFieldName,LatitudeFieldName))

  # Set the coordinate system to GCS Lat\Lon WGS1984
  st_crs(SDF) <- st_crs(4326)
  
  return(SDF)
}


# Counters to increment the table and figure numbers as the report progresses.
TableCount = 1
FigureCount = 1

```


Scott D. Miller  
Information Technology Specialist/Data Manager  
National Park Service, Arctic Inventory and Monitoring Network  
240 W. 5th Ave.  
Anchorage, AK 99501

Kyle Cutting  
Wildlife Biologist  
National Park Service, `r WRST`  
8 Richardson Hwy Mile 106  
Copper Center, AK 99573  

[Add other contributors here]


# Introduction

```{r echo=FALSE}

# Get the summarized data for the comp count survey into a data frame
Sql = paste("SELECT        [Survey name],[Year], Herd, [Survey type], [Type of survey], Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull_Calculated AS Bull, Adult_Calculated AS Adult, Caribou_Calculated AS Caribou, Unknown, [Calf, male], [Calf, female], 
                         [Calf, unclassified], GroupSize_Mean, [Pct. calves], [Pct. cows], [Pct. bull (composition survey)] AS [Pct. bull], [Calves/100 cows], [Bulls/100 cows], [Total groups observed], [Groups in the survey area], 
                         [Groups out of the survey area], [Marked groups], [Marked groups seen], [Marked groups not seen], [Percent marked groups seen], [Percent marked groups not seen], [Total collared animals observed], 
                         [Number of frequencies that were available on survey date (Animal_Movement)], [Collars not heard or not searched], [Total groups marked seen], [Start date], [End date], [Survey days], [Search areas], Season, 
                         CertificationLevel
FROM            Summary_Census
WHERE ([Survey name] = '",SurveyName,"')",sep="")
CompDF = GetDataFrame(Sql)

# Basic survey details
StartDate = CompDF[1,]$`Start date`
EndDate = CompDF[1,]$`End date`
SurveyDays = CompDF[1,]$`Survey days`

```

```{r echo=FALSE}
# Get the historical length of the various surveys into a data frame
Sql = paste("SELECT        Herd, [Survey type], Earliest, [Latest survey], [Years of record], [Survey days], Surveys 
FROM Summary_LengthOfRecordByHerdAndSurveyType Where Herd = '",Herd,"' And [Survey type] = 'Composition count'",sep="")
DF = GetDataFrame(Sql)
if(nrow(DF) > 0){
  Earliest = DF[1,]$Earliest
  TotalSurveys = DF[1,]$Surveys
}
```

Caribou are an integral component of the ecosystem in `r WRST` (WRST) and are of great importance to the ecosystem and to people for both consumptive and non-consumptive uses. WRST is mandated by the Alaska National Interest Lands Conservation Act to protect habitat for caribou, to provide the opportunity for continued subsistence uses, and to conserve natural and healthy wildlife populations. In order to meet these requirements the Park monitors the `r Herd` caribou herd with the goal of providing Park managers with reliable scientific information on the herd's abundance, composition and distribution. 
WRST conducted a herd composition survey from `r StartDate` to `r EndDate` to assess calf recruitment, adult sex ratios, and seasonal distribution. In this report we provide provisional survey results in the form of data and summary statistics, but leave in-depth statistical analysis of population size and composition for a later date.

# Methods

Monitoring of caribou populations in CAKN employs the use of radiocollars and radiotelemetry to locate groups and to provide a mark/recapture estimate of population size and age and sex composition. The use of radiotelemetry is standard throughout Alaska and parts of Canada for monitoring caribou populations.
A sample of 30-40 radiocollared cows is maintained for population assessment. Population assessment is made in two efforts: a post-calving (June) census, when cows are grouped and calf production can be determined; and a fall composition count when bulls associate with cows during the rut. With the mark-recapture estimate of cows from the spring survey and the herd composition obtained from the fall count, the herd size, composition, and calf recruitment can be estimated. Both annual surveys are conducted by aerial observation from a helicopter and small airplane. These caribou monitoring methods were formalized in a collaboration between the WRST and the Central Alaska Inventory and Monitoring Program in 2018 (Putera and Miller, 2018). The WRST caribou monitoring protocol, and standard operating procedures, as well as reports, journal articles and data are available through the NPS's Integrated Resource Management Applications portal at [https://irma.nps.gov/DataStore/Reference/Profile/2217176](https://irma.nps.gov/DataStore/Reference/Profile/2217176). 

# Results

The composition survey took place from `r StartDate` to `r EndDate` over `r SurveyDays` days (Table `r TableCount`). Survey conditions were [fill in survey conditions if available.]. 

Table `r TableCount`. Survey dates and observer/pilot pairs, `r Year`.

```{r echo=FALSE}
# Get the survey days and observers into a frame
Sql = paste("SELECT  DISTINCT CONVERT(Date, SightingDate) AS [Date],   SearchArea, Observer1 AS Observer, Pilot
FROM            CaribouGroups
WHERE        (SurveyName = '",SurveyName,"') AND (SurveyType = 'CC')
ORDER BY [Date],SearchArea, Observer, Pilot",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)
```


## Herd distribution

Caribou were distributed as shown in Figure `r FigureCount`.

```{r echo=FALSE}

# Get the spatial data for comp and population surveys
Sql = paste("SELECT SurveyName,SearchArea, GroupNumber, Lat, Lon,Lat as Latitude,Lon as Longitude FROM Dataset_Census WHERE SurveyName = '",SurveyName,"' And Lat > 0",sep="")
CCDF = GetDataFrame(Sql)

# Convert the data into a spatial data frame
# Get rid of rows with no spatial coordinates - they will cause the map to bomb
WithCoordinatesDF = subset(CCDF,Lat != "NULL" & Lon != "NULL")
CCSDF = GetPointSpatialDataFrameFromDataFrame(WithCoordinatesDF,"Lat","Lon")

# Read the shapefile into a spatial features dataset
# Script users will have to point to a WRST park layer in geographic coordinate system or ask me and I'll send this one
AKParksShp <- sf::read_sf("C:/Work/GIS Common Layers/GCS/AKParks_reprojected.shp")

# Show a map of where caribou groups were found during the survey. Generate a density heat map also.
#Plot the points over the shapefile (shapefile first, then points)
ggplot() + geom_sf(data = AKParksShp) + 
  geom_sf(data = CCSDF, color='black', size= 2, show.legend = T) +
  # geom_text(data = CCSDF, size = 3, color = 'red',mapping = aes(CCSDF$Longitude,CCSDF$Latitude,label = CCSDF$SurveyName), vjust = -1,hjust = 0,inherit.aes = FALSE) +
  
   # Density layer
    geom_density_2d(data = CCDF, aes(x = Longitude, y = Latitude), colour = "red", alpha = 0.9) +
    scale_fill_brewer() + # This seems to make the colors less wacky, not sure why
    guides(fill = guide_legend(title = "Density of observations")) + # This gives a legend to the density
  
  # Limit the view port to just WRST
  xlim(-146.5,-139) +
  ylim(59.8,62.65) +
  #ggtitle(paste("Composition and population survey observations,",Year,sep="") ,subtitle = "Wrangell-St. Elias National Park and Preserve") +
  xlab('Longitude') + ylab('Latitude')

```

Figure `r FigureCount`. Distribution and density of caribou groups in Wrangell - St. Elias National Park and Preserve and outlying areas during the `r SurveyName`.

```{r echo=FALSE}
# Increment the figure counter
FigureCount = FigureCount + 1
```


## Herd Composition

```{r echo=FALSE}

# Store some key values to be put in the report text further on
Caribou = CompDF[1,]$Caribou
BullsTo100Cows = CompDF[1,]$`Bulls/100 cows`
BullsTo100Cows = round(BullsTo100Cows, digits = 1) 
CalvesTo100Cows = CompDF[1,]$`Calves/100 cows`
CalvesTo100Cows = round(CalvesTo100Cows, digits = 1) 

```

We classified `r Caribou` caribou during the survey. Calf recruitment was `r CalvesTo100Cows` per 100 cows and there were `r BullsTo100Cows` bulls per 100 cows (Table `r TableCount`).

Table `r TableCount`. Composition survey results.
```{r echo=FALSE}
# Display the survey summary data from way up top. Transpose it so it fits better in Word.
kable(t(CompDF))

# Increment the table counter
TableCount = TableCount + 1

```

## Historical Results


Tables `r TableCount` and `r TableCount + 1` provide the most recent survey data compared to the results of past composition count surveys for the `r Herd` herd.

### Counts

Table `r TableCount`. Historical composition of the Chisana herd.

```{r echo=FALSE}

# Grab all the historical comp count data for the herd into a frame and show it
# Exclude raw data - provisional or certified only
Sql = paste("SELECT [Year], [Survey type], Cow, Calf, [Small bull], [Medium bull], [Large bull], Bull_Calculated AS Bull, Adult_Calculated AS Adult, Caribou_Calculated AS Caribou, Unknown, [Total groups observed] as n
,CASE WHEN CertificationLevel = 'Certified' THEN '' ELSE '*' END As [Qual.]
FROM            Summary_Census
WHERE Herd = '",Herd,"' 
And [Survey type] = 'CC' 
And CertificationLevel <> 'Raw'
And Year < ",Year,"
ORDER BY Year DESC,Season DESC",sep="")
ComparisonDF = GetDataFrame(Sql)
kable(ComparisonDF)
TableCount = TableCount + 1
```

Survey types: CC = Composition count, PE = Population estimate
Qual.: * = Provisional data

### Composition

Table `r TableCount` and Figures `r FigureCount` and `r FigureCount + 1` summarize historical sex and age ratios and herd composition statistics over time for the `r Herd` herd.

Table `r TableCount`. Historical sex and age ratios and herd composition statistict for the `r Herd` herd.

```{r echo=FALSE}
# Grab all the historical comp count statistical data for the herd into a frame and show it (the table above shows counts; this table shows calculated stats)
# Exclude raw data - provisional or certified only
Sql = paste("SELECT Year, Season,[Survey type], GroupSize_Mean AS [Mean group size], [Pct. calves], [Pct. cows], [Pct. bull (composition survey)] AS [Pct. bull], [Calves/100 cows], [Bulls/100 cows] AS [Bulls/100 cows], 
[Total groups observed] AS n
FROM            Summary_Census
WHERE Herd = '",Herd,"' 
And CertificationLevel <> 'Raw'
And Year < ",Year,"
ORDER BY Year DESC",sep="")
ComparisonDF = GetDataFrame(Sql)

kable(ComparisonDF)

TableCount = TableCount + 1
```

Figure `r FigureCount`. Calves/100 cows over time.
```{r echo=FALSE}

# Plot the calf:cow ratio
ggplot(data=ComparisonDF, aes(x=Year, y=`Calves/100 cows`, group=1)) +
  geom_line()+
  geom_point()

FigureCount = FigureCount + 1
```

Figure `r FigureCount`. Bulls/100 cows over time.
```{r echo=FALSE}

# Plot the bull: cow ratio data
ggplot(data=ComparisonDF, aes(x=Year, y=`Bulls/100 cows`, group=1)) +
  geom_line()+
  geom_point()

FigureCount = FigureCount + 1
```

<!-- # Sightability DOES NOT APPEAR TO BE A BIG PART OF COMP SURVEYS SO COMMENTED OUT -->


<!-- Stats -->

<!-- ```{r echo=FALSE} -->
<!-- Sql = paste("SELECT Year,Season,[Total groups observed] as n -->
<!-- -- , [Groups in the survey area], [Groups out of the survey area] -- the [In] attribute isn't used much -->
<!-- , [Marked groups], [Marked groups seen] -->
<!-- --, [Marked groups not seen] -->
<!-- , [Percent marked groups seen] -->
<!-- --, [Percent marked groups not seen] -->
<!-- --, [Total collared animals observed], [Number of frequencies that were available on survey date (Animal_Movement)] -->
<!-- , [Collars not heard or not searched],  -->
<!--  [Total groups marked seen] -->
<!-- FROM            Summary_Census -->
<!-- WHERE Herd = '",Herd,"' And [Survey type] = 'CC'  -->
<!-- ORDER BY Year DESC",sep="") -->
<!-- ComparisonDF = GetDataFrame(Sql) -->

<!-- kable(ComparisonDF) -->

<!-- ``` -->


<!-- Table `r TableCount`. Composition survey results by search area. [DEFINE THE SEARCH AREAS]. -->
```{r echo=FALSE}
# Sql = paste("SELECT        [Survey name], Herd, [Type of survey], SearchArea, Cow, Calf, [Small bull], [Medium bull], [Large bull], [Bull (composition survey)], [Adult (composition survey)], [Caribou (composition survey)], Unknown, [Calf, male], 
#  [Calf, female], [Calf, unclassified], [Calves/100 cows], [Bulls/100 cows (composition survey)], [Total groups observed], [Groups in the survey area], [Groups out of the survey area], [Marked groups], [Marked groups seen], 
#  [Marked groups not seen], [Percent marked groups seen], [Percent marked groups not seen], [Total collared animals observed], [Total groups marked seen], [Start date], [End date], [Survey days], [Search areas], Park, Year
# FROM Summary_CensusBySearchArea
# WHERE ([Survey name] = '",SurveyName,"')
# ORDER BY [Searcharea]",sep="")
# CCDF = GetDataFrame(Sql)
# kable(t(CCDF))
# TableCount = TableCount + 1
```

# Discussion

[Add discussion here]

# References
ANILCA, 1980

# Appendix A: Dataset Certification Summary

Appendix A contains a summary of data quality certification levels and number of records for the `r SurveyName` (current as of `r Sys.Date()`). Definitions follow the summary table.

```{r echo=FALSE}

# This shows how trustworthy the dataset is. See the table of definitions following
Sql = paste("SELECT SurveyName, CertificationLevel, [Number of records]
FROM     Summary_Census_CertificationLevel
WHERE ([SurveyName] = '",SurveyName,"')",sep="")
DF = GetDataFrame(Sql)
kable(DF)
```

CertificationLevel | Description
--- | ---
Raw | Records imported directly from field surveys with no quality control or validation procedures performed.
Provisional | Records that have undergone minimal quality control and/or have been examined by the project leader and are deemed suitable for internal reports but have not been validated against a published summary or report. Some records may never be fully certified due to the lack of validating materials.
Certified | Records that have undergone extensive quality control procedures and have been validated against published reports, journal articles or data summary reports.



# Appendix B: Survey Description and Notes 

Appendix B contains a description of the survey and any notes about the survey as written by the pilot, observer or biologist. These data are available in the WRST_Caribou.Surveys table.

```{r echo=FALSE}
# A description of the survey and/or notes provided by personnel is stored in the database Surveys table.
# These data may be useful for someone to know - add it in this appendix
# Survey description and notes are stored in the Surveys table. See if there is a record for the survey and store 
# the description and notes in variables.
Sql = paste("SELECT SurveyName, Description, Notes FROM Surveys WHERE ([SurveyName] = '",SurveyName,"')",sep="")
DF = GetDataFrame(Sql)
Description = ""
SurveyNotes = ""
if(nrow(DF) > 0){
  Description = DF[1,]$Description
  SurveyNotes = DF[1,]$Notes
}
```

## Survey Description
`r Description`

## Survey Notes
`r SurveyNotes`

# Appendix C: Data Quality Issues

Appendix C Contains known data quality issues that potential users of the dataset should know (current as of `r Sys.Date()`). 

```{r echo=FALSE}

# Data quality issues are inevitable and stored in the Surveys_Issues database table. Dump them out so the user can see what they are dealing with.
Sql = paste("SELECT Surveys_Issues.IsResolved, Surveys_Issues.DateLogged, Surveys_Issues.Author, Surveys_Issues.Issue
FROM  Surveys INNER JOIN
 Surveys_Issues ON Surveys.SurveyName = Surveys_Issues.SurveyName where Surveys.SurveyName = '",SurveyName,"'
ORDER BY Surveys.SurveyName DESC,Surveys_Issues.IsResolved, Surveys_Issues.DateLogged",sep="")
DF = GetDataFrame(Sql)
kable(DF)

```


# Appendix D: Quality Control Checks Results

Current as of `r Sys.Date()`

Appendix D contains the results of quality control checks on the survey dataset. These results are intended to inform NPS scientists about any data quality problems and should be resolved, if possible, before publication.Move unresolvable issues that data consumers should know about to the Surveys_Issues table and publish from there.

## QC-1: Observed/Available Collars Reconciliation
Ideally, every collar that is registered in the NPS Animal Movement database for the `r Herd` monitoring program is searched for during a composition survey. Table `r TableCount` shows the count of frequencies detected during the `r SurveyName` and the number of collars that were available according to the data in the Animal Movement database. It is not critical that the numbers match, only that they reconcile in a logical way. Collars occasionally fail, the terrain may block their signal, the crew ran out of time or encountered bad weather, or collars were are not detected for another reason. Sometimes collar deployments are not terminated in Animal_Movement after a mortality, resulting in an inflated inventory of available collars. The list below is a starting point for reconciling these issues.

Table `r TableCount`. Observed/Available Collars Reconciliation.

```{r echo=FALSE}
Sql = paste("SELECT 
-- Herd, [Survey type], 
SurveyName as [Survey], [Count of frequencies] as [Observed collar frequencies], AvailableCollars as [Available collar frequencies]
FROM     QC_CaribouGroups_FrequenciesCountNotMatchingAvailableFrequenciesInAnimal_Movement WHERE (SurveyName = '",SurveyName,"')",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)

TableCount = TableCount + 1
```

## QC-2: Marked groups/frequency(ies) detected check
An important component of caribou surveys is sightability which is accomplished through a mark-recapture model. Groups are marked via GPS collars. The first survey flight does not have access to GPS telemetry from the collars and searches for caribou blind. The follow up flight is able to find marked groups via telemetry and tags each observed animal group as marked if it had at least one collared animal in it and the group is marked as 'seen' if the blind group found it. This quality control test shows records having a non-null FrequenciesInGroup attribute and its corresponding Marked attribute marked anything but 1.

Table `r TableCount`. Records having non-null FrequenciesInGroup attribute with Marked attribute not equal to 1 (boolean, True). A blank table below indicates a QC check PASS.

```{r echo=FALSE}
Sql = paste("SELECT SurveyName, SightingDate, GroupNumber, Marked, FrequenciesInGroup
FROM QC_CaribouGroups_GroupsWithFrequenciesDetectedButMarkedIsNULL
WHERE (SurveyName = '",SurveyName,"')",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)

TableCount = TableCount + 1
```

## QC-3: Missing sightability information
Quality control check QC-3 extends check QC-2 by providing information about how many survey records have null Seen and Marked attributes. 

Table `r TableCount`. Surveys where sightability data (completely null Seen and Marked attributes for all survey records) were not collected. A blank table below indicates a QC check PASS.

```{r echo=FALSE}
Sql = paste("SELECT SurveyName as Survey, Seen, Marked
FROM            QC_CaribouGroups_SurveysWithNoSightabilityInformation
WHERE (SurveyName = '",SurveyName,"') ORDER BY SurveyName DESC",sep="")
ObsPilotsDF = GetDataFrame(Sql)
kable(ObsPilotsDF)

TableCount = TableCount + 1
```




